---
title: "bio_assembly"
author: "Stefano Larsen"
date: "10/7/2021"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6)
```


```{r}
#library(readxl)
library(tidyverse)
library(reshape2)
library(ggrepel)
library(vegan)
library(iCAMP)
install.packages('NST')
library(NST)
library(data.table)
library(patchwork)
```

# Run the Sloan neutral model to assess fit and m for each year across LB10 (imputed)
```{r}
snm85=
snm(
LB10.imp %>% filter(year=='1985') %>% select(-c(1,2)), simplify = F)$stats

snm85$stats
```



#plot for fun
```{r}
snm85$detail %>% 
  ggplot()+aes(log(p), freq)+geom_point(aes(col=type), alpha=.5)+
  geom_path(aes(log(p), freq.pred), col='grey30')+
  geom_path(aes(log(p), pred.lwr),linetype='dashed', col='grey30')+
  geom_path(aes(log(p), pred.upr),linetype='dashed', col='grey30')
  
  
```
# Run the Sloan neutral model on each year fillin a list (keeping the fit parameters)
```{r}
snm.list=list()

for(y in 1: length(yearID.lb10)){
  snm.list[[y]]=snm(LB10.imp %>% filter(year==yearID.lb10[y]) %>% select(-c(1,2)), simplify = F)$stats
}

snm.list[[3]]

```

# Run the Sloan neutral model on each year fillin a list (keeping the details only)
```{r}
snm.detail.list=list()

for(y in 1: length(yearID.lb10)){
  snm.detail.list[[y]]=snm(LB10.imp %>% filter(year==yearID.lb10[y]) %>% select(-c(1,2)), simplify = F)$detail
}



```

# plotting the neutral fit from the list
```{r}
snm.detail.list[[28]] %>% 
  ggplot()+aes(log(p), freq)+geom_point(aes(col=type), alpha=.9)+
  geom_path(aes(log(p), freq.pred), col='grey30')+
  geom_path(aes(log(p), pred.lwr),linetype='dashed', col='grey30')+
  geom_path(aes(log(p), pred.upr),linetype='dashed', col='grey30')
```




# create a df holding each year neutral model fits
```{r}
neut.fit=cbind.data.frame(year=yearID.lb10,
rbindlist(snm.list))
```
```{r}
neut.fit %>% 
  ggplot()+aes(AIC, AIC.bino, label=year)+geom_label()+geom_point()+geom_abline(slope=1)
```


# extract the *neutral species types*, for counting the freq of classification for each spp
```{r}
neut.spp.type=
 cbind(Species=unlist(lapply(snm.detail.list, rownames)),# extract the rownames from list (ie species id)
      rbindlist(snm.detail.list), #rbind the neutral fit detail
      year=rep(yearID.lb10, lapply(lapply(snm.detail.list, rownames), length)))  # add (cbind) the year by repeating each y for each spp
      

rep(yearID.lb10, lapply(lapply(snm.detail.list, rownames), length))
  


table(neut.spp.type$Species, neut.spp.type$type)

```

# Add the temperature zonation to the neutral species types
# derive the neutral type for each temp group
```{r}
neut.spp.type$temp_zon=mean_spp_temp_zon$quant_temp[match(neut.spp.type$Species, rownames(mean_spp_temp_zon))]

tmp=
table(neut.spp.type$year, neut.spp.type$type, neut.spp.type$temp_zon)

table(neut.spp.type$type, neut.spp.type$temp_zon)


neutral.cold=
tmp[,,1] %>%
  as.matrix() 


neutral.medium=
tmp[,,2] %>%
  as.matrix() 

neutral.warm=
  tmp[,,3] %>%
  as.matrix() 


```



```{r fig.width=5}
cbind.data.frame(above=neutral.cold[,1], 
                 below=neutral.cold[,2],
                 neutral=neutral.cold[,3]) %>% 
  mutate(totab=rowSums(.), year=yearID) %>% 
  ggplot()+aes(year, below/totab, group=1)+geom_path()+geom_smooth()+
  theme(axis.text.x = element_text(angle=45, hjust=1))

cbind.data.frame(above=neutral.cold[,1], 
                 below=neutral.cold[,2],
                 neutral=neutral.cold[,3]) %>% 
  mutate(totab=rowSums(.), year=yearID) %>% 
  ggplot()+aes(year, above/totab, group=1)+geom_path()+geom_smooth()+
  theme(axis.text.x = element_text(angle=45, hjust=1))


cbind.data.frame(above=neutral.cold[,1], 
                 below=neutral.cold[,2],
                 neutral=neutral.cold[,3]) %>% 
  mutate(totab=rowSums(.), year=yearID) %>% 
  ggplot()+aes(year, neutral/totab, group=1)+geom_path()+geom_smooth()

```


```{r}
cbind.data.frame(above=neutral.warm[,1], 
                 below=neutral.warm[,2],
                 neutral=neutral.warm[,3]) %>% 
  mutate(totab=rowSums(.), year=yearID) %>% 
  ggplot()+aes(year, above/totab, group=1)+geom_path()+geom_smooth()


cbind.data.frame(above=neutral.warm[,1], 
                 below=neutral.warm[,2],
                 neutral=neutral.warm[,3]) %>% 
  mutate(totab=rowSums(.), year=yearID) %>% 
  ggplot()+aes(year, below/totab, group=1)+geom_path()+geom_smooth()
```




# Plot the timeseries of params from neutral fit
```{r fig.width=8, fig.height=4}
neut.fit %>% 
  ggplot()+aes(as.numeric(as.character(year)), m) + geom_path() +geom_smooth(se=F)+

  neut.fit %>% 
  ggplot()+aes(as.numeric(as.character(year)), Rsqr) + geom_path() +geom_smooth(se=F)

```
# which is the winter nao data covering the lb10 imputed ?
```{r}
#ts.NAOw$cdat
NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1985:2018))
```

# *Some indication that positive NAOw phases lead to smaller R2 for the neutral model (less neutral metacom)*
```{r}
cbind.data.frame(wNAO=NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1985:2018)) %>% pull(mean.nao),
     NeutRsqr=neut.fit$Rsqr, Year=time.lb10) %>% 
  ggplot()+aes(wNAO, NeutRsqr)+geom_point(aes(col=Year))+geom_smooth(method='lm')+ylab('Fit of neutral model')+scale_color_gradient(high='darkblue', low='gold3')

summary(lm(NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1985:2018)) %>% pull(mean.nao)~
     neut.fit$Rsqr  ))


# No effect of smooth nao
cbind.data.frame(wNAOsmooth=smoothNAOw  %>% filter(winter.year %in% c(1985:2018)) %>% pull(sNAO),
     NeutRsqr=neut.fit$Rsqr, Year=time.lb10) %>% 
  ggplot()+aes(wNAOsmooth, NeutRsqr)+geom_point(aes(col=Year))+geom_smooth(method='lm')+ylab('Fit of neutral model')+scale_color_gradient(high='darkblue', low='gold3')



```


# explore wavelet from the neutral model params

```{r}
t_neut.fit.m=cleandat(neut.fit$m, times= as.numeric(as.character(neut.fit$year)), clev=4)$cdat

t_neut.fit.r2=cleandat(neut.fit$Rsqr, times= as.numeric(as.character(neut.fit$year)), clev=4)$cdat
```

# Plot WT from the m parameter in neutral fit
```{r}
plotmag(wt(t_neut.fit.m, times=as.numeric(as.character(neut.fit$year))), title='Wavelet transform migr.rate')
```

```{r}
plotmag(wt(t_neut.fit.r2, times=as.numeric(as.character(neut.fit$year))), title='Wavelet transform neutral R2')
```


# *Clean NAOw 85-18*

```{r}
ts.NAOw.85_18=
cleandat(NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1985:2018)) %>% pull(mean.nao),
         times=NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1985:2018)) %>% pull(winter.year) %>% as.numeric(),
         clev=4)$cdat
```

# Coherence migration rate vs wNAO
```{r}
coh.NAOw_neut.m=
coh(ts.NAOw.85_18, t_neut.fit.m, times= as.numeric(as.character(neut.fit$year)),  
    norm='powall', sigmethod = 'fast', nrand = 10000)
```

```{r}
plotmag(coh.NAOw_neut.m)

coh.NAOw_neut.m=bandtest(coh.NAOw_neut.m, c(4,5))
coh.NAOw_neut.m=bandtest(coh.NAOw_neut.m, c(6,7))

get_bandp(coh.NAOw_neut.m)
```




```{r}
coh.NAOw_neut.r2=
coh(ts.NAOw.85_18, t_neut.fit.r2, times= as.numeric(as.character(neut.fit$year)),  
    norm='powall', sigmethod = 'fast', nrand = 10000)
```


# *Winter NAO also shows coherence at 4-6 timescales with R2 from neutral model*
```{r}
plotmag(coh.NAOw_neut.r2)


coh.NAOw_neut.r2=bandtest(coh.NAOw_neut.r2, c(3,4))
coh.NAOw_neut.r2=bandtest(coh.NAOw_neut.r2, c(5,6))

get_bandp(coh.NAOw_neut.r2)
```


##### Stochasticity each year


```{r}
cbind.data.frame(Sites=lb10.siteID, type=)
```

# some test on how to perform the tNST on a loop for each year
```{r}

zio=
tNST(comm=LB10.imp %>% filter(year=='1985') %>% select(-c(1,2)) ,
     dist.method="jaccard",
     group =data.frame(landuse=c('AcM','AcM', 'AcM','AcM', 'AcF', 'AcF','AcF', 'NeM','NeM','AcF')) %>% 
            `rownames<-`(LB10.imp %>% filter(year==yearID.lb10[1]) %>% rownames()),
          abundance.weighted=TRUE, rand=10,
          nworker=1, null.model="PF", between.group = T,
          SES=TRUE, RC=TRUE)$index.pair$MST.ij.ruzicka


mean(zio$index.pair.grp$MST.ij.ruzicka)
mean(zio$index.pair$MST.ij.ruzicka)


```


# get stochasticity each year
```{r}
#create empty matrix to hold the mean stochasticity
tNST.year=as.data.frame(matrix(nrow=34, ncol=2))
tNST.year[,1]=yearID.lb10

for(y in 1: length(yearID.lb10)){ # loop over years
  tNST.year[y,2]=
    mean(tNST(comm=LB10.imp %>% filter(year==yearID.lb10[y]) %>% select(-c(1,2)) , # get the mean of MST.ij
     dist.method="jaccard",
     group =data.frame(landuse=c('AcM','AcM', 'AcM','AcM', 'AcF', 'AcF','AcF', 'NeM','NeM','AcF')) %>% # must create a group (such as landuse)
            `rownames<-`(LB10.imp %>% filter(year==yearID.lb10[y]) %>% rownames()),
          abundance.weighted=TRUE, rand=200,
          nworker=1, null.model="PF", between.group = T,
          SES=TRUE, RC=TRUE)$index.pair$MST.ij.ruzicka)
}

names(tNST.year)=c('year', 'mean.MST')
tNST.year$year=as.numeric(as.character(tNST.year$year))
```

# add the wNAO on the stochasticity df
```{r}
tNST.year$wNAO= wNAO=NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1985:2018)) %>% pull(mean.nao)

tNST.year$smoothNAO=smoothNAOw %>% filter (winter.year %in% c(1985:2018) ) %>% pull(sNAO)
```


```{r}
tNST.year %>% 
  ggplot()+aes(year, mean.MST)+geom_path()


tNST.year %>% 
  ggplot()+aes(wNAO, mean.MST)+geom_point()+geom_smooth(method='lm')

tNST.year %>% 
  ggplot()+aes(smoothNAO, mean.MST)+geom_point()+geom_smooth(method='lm')



```

# wavelet analysis of the stochasticity
```{r}
t_MST=cleandat(tNST.year$mean.MST, times= tNST.year$year, clev=4)$cdat
```

```{r}
plotmag(wt(t_MST, times=tNST.year$year))
```
```{r}
coh.NAOw_MST=
coh(ts.NAOw.85_18, t_MST, times= tNST.year$year,  
    norm='powall', sigmethod = 'fast', nrand = 10000)

coh.NAOw_MST=
coh(ts.NAOw.85_18, t_MST, times= tNST.year$year,  
    norm='powall', sigmethod = 'fftsurrog1', nrand = 100)


```

```{r}
plotmag(coh.NAOw_MST)
coh.NAOw_MST=bandtest(coh.NAOw_MST, c(8,9))
```




#### Extract the betadiversit ybased on hill numbers to check synchrony in year-to-year variation for rare and commmon spp
```{r}
install.packages('hillR')
library(hillR)
```


# Steps to extract the year-to-year variation
```{r}

LB10.imp %>% 
  filter(Code=='LI8') %>% select(-c(1,2)) %>% 
  `rownames<-`(time.lb10) %>% 
  hill_taxa_parti_pairwise(q=0) %>% 
  mutate(year1_year2 = paste(site1, site2, sep='_')) %>% 
  filter(year1_year2 %in% (deltamat)) %>% 
  select(TD_beta)
```

# extract the hillR beta q=0
```{r}
hillR_beta=as.data.frame(matrix(nrow=length(time.delta), ncol=3))
rownames(hillR_beta)=deltamat

for(s in 1:length(lb10.siteID)){
  hillR_beta[,s]=LB10.imp %>% 
  filter(Code==lb10.siteID[s]) %>% select(-c(1,2)) %>% 
  `rownames<-`(time.lb10) %>% 
  hill_taxa_parti_pairwise(q=0) %>% 
  mutate(year1_year2 = paste(site1, site2, sep='_')) %>% 
  filter(year1_year2 %in% (deltamat)) %>% 
  select(TD_beta)
}

names(hillR_beta)=lb10.siteID

hillR_beta.q0=hillR_beta
rm(hillR_beta)

```

# extract hillR beta for q=1
```{r}
hillR_beta.q1=as.data.frame(matrix(nrow=length(time.delta), ncol=3))
rownames(hillR_beta.q1)=deltamat

for(s in 1:length(lb10.siteID)){
  hillR_beta.q1[,s]=LB10.imp %>% 
  filter(Code==lb10.siteID[s]) %>% select(-c(1,2)) %>% 
  `rownames<-`(time.lb10) %>% 
  hill_taxa_parti_pairwise(q=1) %>% 
  mutate(year1_year2 = paste(site1, site2, sep='_')) %>% 
  filter(year1_year2 %in% (deltamat)) %>% 
  select(TD_beta)
}

names(hillR_beta.q1)=lb10.siteID
```

# extract hillR beta for q=2
```{r}
hillR_beta.q2=as.data.frame(matrix(nrow=length(time.delta), ncol=3))
rownames(hillR_beta.q2)=deltamat

for(s in 1:length(lb10.siteID)){
  hillR_beta.q2[,s]=LB10.imp %>% 
  filter(Code==lb10.siteID[s]) %>% select(-c(1,2)) %>% 
  `rownames<-`(time.lb10) %>% 
  hill_taxa_parti_pairwise(q=2) %>% 
  mutate(year1_year2 = paste(site1, site2, sep='_')) %>% 
  filter(year1_year2 %in% (deltamat)) %>% 
  select(TD_beta)
}

names(hillR_beta.q2)=lb10.siteID
```


# clean the hillR beta timeseries
```{r}
t_hillR_beta.q0=cleandat(t(hillR_beta.q0), times=time.delta, clev=4)$cdat
t_hillR_beta.q1=cleandat(t(hillR_beta.q1), times=time.delta, clev=4)$cdat
t_hillR_beta.q2=cleandat(t(hillR_beta.q2), times=time.delta, clev=4)$cdat
```

# phasor hillR q=0
*similar to the overall synch in Bray curtis variation, and richness synchrony*
```{r}
plotmag(
  wpmf(t_hillR_beta.q0, times=time.delta, sigmethod = 'quick', nrand = 1000), title='hillR beta q=0'
)
```

# phasor hillR q=1
```{r}
plotmag(
  wpmf(t_hillR_beta.q1, times=time.delta, sigmethod = 'quick', nrand = 1000),title='hillR beta q=1'
)
```
# phasor hillR q=2
```{r}
plotmag(
  wpmf(t_hillR_beta.q2, times=time.delta, sigmethod = 'quick', nrand = 1000),title='hillR beta q=2'
)
```
### examine coherence with wNAO
#*no coherence with wANO for hillR beta q0*
```{r}
coh_wNAO_betaq0=
  coh(t_NAOw_86.18.exp, t_hillR_beta.q0, time.delta,  norm='powall', sigmethod = 'fast', nrand = 10000)


plotmag(coh_wNAO_betaq0)
```


```{r}
coh_wNAO_betaq1=
  coh(t_NAOw_86.18.exp, t_hillR_beta.q1, time.delta,  norm='powall', sigmethod = 'fast', nrand = 10000)


plotmag(coh_wNAO_betaq1)
```
# *some coherence between wNAO and hillR beta q=2*
```{r}
coh_wNAO_betaq2=
  coh(t_NAOw_86.18.exp, t_hillR_beta.q2, time.delta,  norm='powall', sigmethod = 'fast', nrand = 10000)

coh_wNAO_betaq2=bandtest(coh_wNAO_betaq2, c(4,5))
plotmag(coh_wNAO_betaq2)
```



```{r}
cbind(hillR_beta.q0, time.delta,  wNAO=NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1986:2018)) %>% pull(mean.nao)) %>% 
  as.data.frame() %>% 
  pivot_longer(cols=c(1:10)) %>% 
  ggplot()+aes(wNAO, value)+geom_point(aes(col=time.delta))+scale_color_gradient(high='darkblue', low='gold3')+
  geom_smooth(method='lm')+facet_wrap(~name)


cbind(hillR_beta.q2, time.delta,  wNAO=NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1986:2018)) %>% pull(mean.nao)) %>% 
  as.data.frame() %>% 
  pivot_longer(cols=c(1:10)) %>% 
  ggplot()+aes(wNAO, value)+geom_point(aes(col=time.delta))+scale_color_gradient(high='darkblue', low='gold3')+
  geom_smooth(method='lm')+facet_wrap(~name)


```

# Plotting the year-to-year variation based on hill betadiv
*we see strong effects of some years, especially on beta q=2, the dominant taxa*
```{r fig.width=9, fig.height=6}
hillR_beta.q0 %>% 
  mutate(time.delta=time.delta) %>% 
  mutate(deltamat=deltamat) %>% 
  pivot_longer(cols=c(1:10)) %>% 
  ggplot()+aes(time.delta, value, group=1)+geom_path()+facet_wrap(~name)+theme(axis.text.x = element_text(angle = 90, size=6))+ylab('Beta q=0')

hillR_beta.q2 %>% 
  mutate(time.delta=time.delta) %>% 
  mutate(deltamat=deltamat) %>% 
  pivot_longer(cols=c(1:10)) %>% 
  ggplot()+aes(time.delta, value, group=1)+geom_path()+facet_wrap(~name)+theme(axis.text.x = element_text(angle = 90, size=6))+ylab('Beta q=2')

```

# *overall the influence of wNAO seems stronger on variation of dominant species rather than rare, as seen from beta q=0 and q=2* 
```{r}
cbind.data.frame(Beta.q0.means=rowMeans(hillR_beta.q0), time.delta,  wNAO=NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1986:2018)) %>% pull(mean.nao)) %>% 
  #as.data.frame() %>% 
  #pivot_longer(cols=c(1:10)) %>% 
  ggplot()+aes(wNAO, Beta.q0.means)+geom_point(aes(col=time.delta))+scale_color_gradient(high='darkblue', low='gold3')+
  geom_smooth(method='lm')

cbind.data.frame(Beta.q1.means=rowMeans(hillR_beta.q1), time.delta,  wNAO=NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1986:2018)) %>% pull(mean.nao)) %>% 
  #as.data.frame() %>% 
  #pivot_longer(cols=c(1:10)) %>% 
  ggplot()+aes(wNAO, Beta.q1.means)+geom_point(aes(col=time.delta))+scale_color_gradient(high='darkblue', low='gold3')+
  geom_smooth(method='lm')



cbind.data.frame(Beta.q2.means=rowMeans(hillR_beta.q2), time.delta,  wNAO=NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1986:2018)) %>% pull(mean.nao)) %>% 
  #as.data.frame() %>% 
  #pivot_longer(cols=c(1:10)) %>% 
  ggplot()+aes(wNAO, Beta.q2.means)+geom_point(aes(col=time.delta))+scale_color_gradient(high='darkblue', low='gold3')+
  geom_smooth(method='lm')
```



#################################################################
#Test with spatial beta-diversity change over each year, using Tatsumi decomposition 

#The ecopart.multi function, from Tatsumi (ecol.lett)
This function breaks down the contribution of colinization and extirpation year-to-year to the changes in spatial beta at the 
metacomm level, and whether they lead to homogenization or heterogen of spatial beta
```{r}
ecopart.multi <- function(d1, d2, part="two"){
  part <- match.arg(part, c("two", "four", "sp"))
  N <- nrow(d1) # Number of sites (communities)
S <- ncol(d1) # Number of species
# Delta subscripts (δxyz)
X <- colSums(d1)
Y <- colSums(d2)
Z <- sapply(1:S, function(s) sum(d1[,s]==1 & d2[,s]==1))
# Whittaker's beta and a scaling factor (H)
Alpha1 <- mean(rowSums(d1))
Alpha2 <- mean(rowSums(d2))
Gamma1 <- sum(colSums(d1) > 0)
Beta1 <- Gamma1/Alpha1
H <- Beta1 / ((Alpha2-Alpha1)/Alpha1 + 1)
# Components of temporal changes in beta diversity (ΔΒ)
# See Eq. 4 in the main text
DBeta <- matrix(nrow=4, ncol=S)
DBeta[1,] <- (X *H/N/Alpha1 - H/Gamma1) * (X>0 & Z==0)
DBeta[2,] <- (X-Z) *H/N/Alpha1 * (X>Z & Z>0)
DBeta[3,] <- (-Y *H/N/Alpha1 + H/Gamma1) * (Y>0 & Z==0)
DBeta[4,] <- (Z-Y) *H/N/Alpha1 * (Y>Z & Z>0)
# Store ΔΒ in a vector or matrix
if(part=="two"){
  Res <- c(sum(DBeta[1:2,]), sum(DBeta[3:4,]))
  names(Res) <- c("extirpation", "colonisation")
} else if(part=="four"){
  Res <- c(sum(DBeta[1, DBeta[1,]<0]),
           sum(DBeta[1, DBeta[1,]>0]) + sum(DBeta[2,]),
           sum(DBeta[3, DBeta[3,]>0]),
           sum(DBeta[3, DBeta[3,]<0]) + sum(DBeta[4,]))
  names(Res) <- c("ext.hmgn", "ext.htrg", "col.htrg", "col.hmgn")
} else if(part=="sp"){
  Res <- rbind(sapply(1:S, function(s) sum(DBeta[1:2, s])),
               sapply(1:S, function(s) sum(DBeta[3:4, s])))
  rownames(Res) <- c("extirpation", "colonisation")
  colnames(Res) <- colnames(d1)
}
# Return the vector or matrix
return(Res)
}
```



# The funciton needs two df from two years
return four measures of betadiv change, with negative being decline in betadiv (homogeniz)
```{r}
y1992=
LB10.imp %>% 
  filter(year==1992) %>% select(-c(1,2))

y1993=
  LB10.imp %>% 
  filter(year==1993) %>% select(-c(1,2))

ecopart.multi(y1992, y1993, part='four')

```

```{r}
ecopart.df=
  as.data.frame(matrix(nrow=length(time.delta), ncol=4))
 rownames(ecopart.df)=time.delta

 
 for(y in 1: length(time.delta)){
   ecopart.df[y,]=
     
     ecopart.multi(LB10.imp %>% filter(year==yearID.lb10[y]) %>% select(-c(1,2)),
                   LB10.imp %>% filter(year==yearID.lb10[y+1]) %>% select(-c(1,2)), part='four' )
 }
 
 
names(ecopart.df)= c('ext.hmgn',    'ext.htrg'  ,  'col.htrg'   , 'col.hmgn' ) 

ecopart.df$delta.Whittbeta=rowSums(ecopart.df)
ecopart.df$time.delta=deltamat
```

```{r fig.width=8}
ecopart.df %>% 
  pivot_longer(cols=c(1:5)) %>% 
  ggplot()+aes(time.delta, value, group=name)+geom_path(aes(col=name), size=1)+theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  scale_color_manual(values=c('#4F90EE', '#2F60E3', 'grey70', '#E3682F', '#CD480B'))+geom_hline(yintercept = 0, col= 'grey50', linetype='dashed')+  ylab('Change in spatial beta-diversity')+xlab(NULL)
```



# this section stems from the LB_trait script, where the trait dist matrix is used to run iCAMP



# how many missing trait info?
```{r}
missing.trt=
apply(Spp_subs_trt, 1, function(x) sum(is.na(x)))
```

# derive spp distance matrix based on traits (excluding spp with n missing values)
```{r}
trt.dist=
vegdist(Spp_subs_trt [names(missing.trt[missing.trt< 45]),], na.rm=T)



sum(is.na(trt.dist))

min(zio)
max(zio)
```

```{r}
library(iCAMP)
```


# select only spp with  valid functional dist values
```{r}
xx=LB10.imp %>% select(names(missing.trt[missing.trt< 45]))
```

# first test with assembly analysis using iCAMP and the trait distance matrix
# *run the iCAMP 'qpen' over each year*
```{r}
icamplist=list()

for(y in 1: length(yearID.lb10)){
  icamplist[[y]]= qpen(
LB10.imp %>% filter(year==yearID.lb10[y]) %>% select(names(missing.trt[missing.trt< 45])),
as.matrix(trt.dist), rand.time = 200)
}



```

```{r}
icamplist[[1]]$ratio

icamplist.ratio=list()
for(y in 1:34){icamplist.ratio[[y]]=icamplist[[y]]$ratio}
```

```{r}
assembly_ratios=
rbindlist(icamplist.ratio)

assembly_ratios$year=yearID.lb10
assembly_ratios$wNAO=wNAO
```
# *No evident patterns in assembly processes over time or vs NAO*
```{r}
assembly_ratios %>% 
  ggplot()+aes(as.numeric(as.character(year)), Dispersal.Limitation)+geom_smooth()

assembly_ratios %>% 
  ggplot()+aes(as.numeric(as.character(year)), Homogenizing.Dispersal)+geom_smooth()



assembly_ratios %>% 
  ggplot()+aes(wNAO, Dispersal.Limitation)+geom_point()


assembly_ratios %>% 
  ggplot()+aes(wNAO, Undominated)+geom_point()+geom_smooth(method = 'lm')

assembly_ratios %>% 
  ggplot()+aes(wNAO, Homogenizing.Dispersal)+geom_point()+geom_smooth(method = 'lm')
```


