---
title: "GCB_SpatialSynchrony"
author: "Stefano Larsen"
date: "2023-08-16"
output: html_document
---


```{r}

install.packages('ggrepel')
library(readxl)
library(tidyverse)
library(reshape2)
library(ggrepel)
library(zoo)
library(vegan)
library(codyn)
library(iCAMP)
library(data.table)
install.packages('gdata')
library(gdata)
library(corrplot)
library(RColorBrewer)
```



```{r}
theme_set(theme_bw())
```



###### ###################
### *Biotic synchrony* ###

########################


# Import the invertebrate data as given by Isabelle
*will work on subsetting, species, sites, and imputing some missing species abundance values*
```{r}
bugs<- read.csv("~/Documents/LB_synchrony/Brianne Inverts to 2018 .csv")
```

# wotk on the site code. Add 'I' and remove 'DLB' from the string names
```{r}
bugs$Code=str_remove(bugs$Code, "DLB")
str_sub(bugs$Code, 2,1)='I'
bugs$Code
bugs$X=NULL
```

# How many years per stream
*site.occur* 
```{r}
table(bugs$Code, bugs$year)
rowSums(table(bugs$Code, bugs$year))

site.occur=(table(bugs$Code, bugs$year))

```

# Plot the site occurrence, number of years sampled
*need to impute missing years for each spp in some sites*
# year 
```{r fig.width=6}
site.occur %>% 
  as.data.frame() %>% 
  #filter(Var2!=c('1981', '1982')) %>% 
 ggplot()+aes(Var2, Var1)+geom_tile(col='grey40',aes(fill=as.factor(Freq)))+xlab(NULL)+ylab(NULL)+theme(legend.position = 'none')+
  theme(axis.text.x = element_text(angle = 90, vjust =0.4, hjust =-1))
```

# To create a complete and common df I should exclude LI3,LI5,GI1,GI2,CI6,CI3 and
1981,1982,1994, 2008,2009,2011


```{r}
site.occur %>% 
  as.data.frame() %>% 
  filter(!Var2 %in% c('1981', '1982', '1994', '2008','2009','2011')) %>%
  filter(!Var1 %in% c( 'LI3','LI5','GI1','GI2','CI6','CI3')) %>% 
 ggplot()+aes(Var2, Var1)+geom_tile(col='grey40',aes(fill=as.factor(Freq)))+theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```
# Look at the LB10 sites with actual years holes, to be filled

```{r}
site.occur %>% 
  as.data.frame() %>% 
  filter(!Var2 %in% c('1981', '1982')) %>%
  filter(!Var1 %in% c( 'LI3','LI5','GI1','GI2','CI6','CI3')) %>% 
 ggplot()+aes(Var2, Var1)+geom_tile(col='grey40',aes(fill=as.factor(Freq)))+theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
 ggtitle('LB10 most complete')+xlab(NULL)+ylab(NULL)
```





# overall spp abundance and occurrences
```{r fig.width=7}
spp_occur=sort(colSums((decostand(bugs[,-c(1,2)], 'pa'))), decreasing = T)
plot(spp_occur)

spp_abb=sort(colSums(bugs[,-c(1,2)]), decreasing = T)
plot(spp_abb)

plot(log(spp_abb), log(spp_occur))

```


```{r}
spp_site_occur=
  sort(
  bugs %>% 
  select(-year) %>% 
  group_by(Code) %>% summarise_all(sum) %>% 
  select(-Code) %>% 
  decostand('pa') %>% 
  colSums(),
  decreasing = T)

```


```{r}
spp_year_occ=
bugs %>% 
  select(-Code) %>% 
  group_by(year) %>% summarise_all(sum) %>% 
  select(-year) %>% 
  decostand('pa') %>% 
  colSums() %>% 
  sort(decreasing = T)
```




# gather bugs data in long format -useful for subsequent manipulations
*this still contains all spp and sites*
```{r}
bugs.long=
bugs %>% 
  gather(Species, abund, -c(Code, year))
```



# Here subsetting *spp occurring in > 5 streams & more than 7years*, while excluding certain streams for good (too few years).
# The bugs subset with species occurring in > 5 streams *bugs.subs6*
# LI5 seems to have many years, but large hole in 1995-2000
# this is also like *LB10* in Ecology paper
```{r}

bugs.subs6=
  cbind(bugs %>%
        filter(!year %in% c('1981', '1982', '1994', '2008','2009','2011')) %>% # excluding these years
        filter(!Code %in% c( 'LI3','LI5','GI1','GI2','CI6','CI3')) %>% # excluding these sites
        select(Code,year) ,
  
        bugs %>% 
        filter(!year %in% c('1981', '1982', '1994', '2008','2009','2011')) %>%
        filter(!Code %in% c( 'LI3','LI5','GI1','GI2','CI6','CI3')) %>% 
        select( names(spp_site_occur[spp_site_occur>5]) & names(spp_year_occ[spp_year_occ>7])) 
       )
```



# Create a vector with species identity for the reduced df with 10 streams and no holes
```{r}
speciesID.subs=
colnames(bugs.subs6[,-c(1,2)])
```

# The list of *species synchrony* for each species between each stream in the reduced set of sites (LB10- bugs.subs6). Based on simple correlation
*still with a few holes (no species imputation yet)*
```{r}
spp.list.corr<-list()

for(spp in 1:length(speciesID.subs)) {
  spp.list.corr[[spp]]= 
  #speciesID.subs=as.factor(names(spp_site_occur[spp_site_occur>6]))
  
  sppcorr=
  bugs.subs6 %>% 
  select(Code, year, speciesID.subs[spp]) %>%  
  spread(Code, value=speciesID.subs[spp]) %>% select(-year) %>% 
  cor(method='spearman') %>% 
  gdata::unmatrix() %>% 
  data.frame() %>% 
  cbind(rep(speciesID.subs[spp], 1 )) %>% 
    cbind(row.names(.)) %>% 
  `colnames<-`(c('synch', 'Species','site.pair'))# use this function: `colnames<-` to assign column names within a pipe
    
  }
  

spp.list.corr[[17]]

```

# Create a list similar to spp.list.corr but as simple correlation matrix. Used for calculating the mean across species.

```{r}
list2<-list()

for(spp in 1:length(speciesID.subs)) {
  list2[[spp]]= 
  #speciesID.subs=as.factor(names(spp_site_occur[spp_site_occur>6]))
  
  sppcorr=
  bugs.subs6 %>% 
  select(Code, year, speciesID.subs[spp]) %>%  
  spread(Code, value=speciesID.subs[spp]) %>% select(-year) %>% 
  cor(method='spearman')
}

list2[41]
```

# Derive the mean synchrony across sites based on the mean synchrony of species in LB10 
```{r}
array2=array(as.numeric(unlist(list2)), dim=c(length(unique(bugs.subs6$Code)), length(unique(bugs.subs6$Code)), length(speciesID.subs)))
dimnames(array2)=list( unique(bugs.subs6$Code), unique(bugs.subs6$Code),speciesID.subs )

array2[,,1]

```

# The mean synchrony across LB10 sites based on simple correlation across populations from species occurring > 6 streams
```{r}
mean_spp_corr.mat=
apply(array2, c(1,2), mean, na.rm=T)
```

```{r fig.width=7}
install.packages('corrplot')
library(corrplot)
corrplot(round(mean_spp_corr.mat,2), method='color', type='lower', addCoef.col = "black", title = 'Mean pop synchrony (corr)', mar=c(0,0,2,0), number.cex = .8, tl.cex = 0.8)
```




# Combine the population synchronies in the list in one long synchronies dataframe
#*synchronies_spp*
```{r}
synchronies_spp=
rbindlist(spp.list.corr)

synchronies_spp=
synchronies_spp %>% 
  filter(synch != 1)
```
# add the *stream distances*
```{r}
synchronies_spp$dist=dist.pairs$dist[match(synchronies_spp$site.pair, dist.pairs$site_pairs)]
```


#add the mean species synchrony overall
```{r}
synchronies_spp %>% 
  group_by(Species) %>% 
  mutate(mean.spp.synch=mean(synch))
```




# Add the columns of 'from' and 'to' in the synchronies_spp df
```{r}

synchronies_spp=cbind(synchronies_spp,
colsplit(synchronies_spp$site.pair, ':', c('from', 'to'))
                     )
```
# Add the chemical distance (derived in the 'Env.data' script) to the species synchrony df
```{r}
synchronies_spp$chem_dist=synchronies.df$chem.dist[match(synchronies_spp$site.pair, synchronies.df$site.pair2)]

#ad the chemical summer synch
synchronies_spp$chem_synch.s=synchronies.df$synch.s [match(synchronies_spp$site.pair, synchronies.df$site.pair2)]

# add the chamical winter synch
synchronies_spp$chem_synch.w=synchronies.df$synch.w [match(synchronies_spp$site.pair, synchronies.df$site.pair2)]

```




# Plot the mean populations synchrony vs chemical distance...
```{r}
synchronies_spp %>% 
  ggplot()+aes(chem_dist, synch)+geom_smooth()+theme_bw()+
  geom_point(aes(col=Species), alpha=0.2)+theme(legend.position = 'none')+
  ylab('Population synchrony (corr)')
```


# Plot the mean populations synchrony vs chemical synchr...
*no apparent relation between simple population synch and chemical (PC1) simple synchrony*
```{r}
synchronies_spp %>% 
  ggplot()+aes(chem_synch.s, synch)+geom_smooth()+theme_bw()+
  geom_point(aes(col=Species), alpha=0.2)+theme(legend.position = 'none')+
  ylab('Population synchrony (corr)')

synchronies_spp %>% 
  ggplot()+aes(chem_synch.w, synch)+geom_smooth()+theme_bw()+
  geom_point(aes(col=Species), alpha=0.2)+theme(legend.position = 'none')+
  ylab('Population synchrony (corr)')


synchronies_spp %>% 
  ggplot()+aes(chem_synch.s, synch)+geom_smooth()+theme_bw()+
  theme(legend.position = 'none')+
  ylab('Population synchrony (corr)')

synchronies_spp %>% 
  ggplot()+aes(chem_synch.w, synch)+geom_smooth()+theme_bw()+
  theme(legend.position = 'none')+
  ylab('Population synchrony (corr)')


```





#######################################
## Imputation of missing species data
###########################


# Look at the subsetted LB10 data 
*again selecting (as with bugs.subs6) spp occurring in >5 streams and more than 7y* 
```{r}


bugs.imp=
 cbind(bugs %>%
        filter(!year %in% c('1981', '1982')) %>% # excluding these years
        filter(!Code %in% c( 'LI3','LI5','GI1','GI2','CI6','CI3')) %>% # excluding these sites
        select(Code,year) ,

bugs %>% 
        filter(!year %in% c('1981', '1982')) %>%
        filter(!Code %in% c( 'LI3','LI5','GI1','GI2','CI6','CI3')) %>% 
        select( names(spp_site_occur[spp_site_occur>5]) & names(spp_year_occ[spp_year_occ>7]))
)
```


 Which sites in LB10?
```{r fig.width=6}
site.occur %>% 
  as.data.frame() %>% 
  filter(!Var2 %in% c('1981', '1982')) %>%
  filter(!Var1 %in% c( 'LI3','LI5','GI1','GI2','CI6','CI3')) %>% 
 ggplot()+aes(Var2, Var1)+geom_tile(col='grey40',aes(fill=as.factor(Freq)))+theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
 ggtitle('LB10 most complete')+theme(legend.position = 'none')+xlab(NULL)+ylab(NULL)
```

# Load some packages for imputation functions
```{r}
install.packages('imputeTS')
library(imputeTS)

install.packages('zoo')
library(zoo)
```



# Function that uses the *imputeTS::na_ma function* to get the moving average (4y before & 4y after) mean for missing values (using floor to convert <1 to zeros)
```{r}
NA2ma=function(x){
  x=floor(imputeTS::na_ma(x, 4, weighting='simple'))
  return(x)
}


NA2ma(c(2,3,5,6,7,81,NA,6,6,6,6))
```




# isolate single site for *imputation for each species*
# *CI1.subs*
*imputation using the subset data with spp in >5 sites*
```{r}
b.CI1.imp=
  bugs.imp %>% 
  filter(Code=='CI1')

b.CI1.imp=# add extra row with missing year and NA for spp abundances
b.CI1.imp %>% 
  add_row(Code='CI1', year=1991)

# sort to order by year
b.CI1.imp=
b.CI1.imp[order(b.CI1.imp$year),]


# impute species as median abundance for the entire time series
b.CI1.imp=
  cbind.data.frame(b.CI1.imp[,c(1,2)],
   apply(b.CI1.imp[,-c(1,2)], 2, NA2ma)
  )

# just a comparison of the effect of imputing using the overall sp median (make too low imputation)
zio=
  cbind.data.frame(b.CI1.subs[,c(1,2)],
   apply(b.CI1.subs[,-c(1,2)], 2, NA2median)
  )

```

```{r}
plot(
rowSums(decostand(b.CI1.imp[,-c(1,2)], 'pa')), type='l')


plot(
rowSums(decostand(zio[,-c(1,2)], 'pa')), type='l')

```

# CI2 imputation
*this site is missing 1991 (as all sites) and 2009 , 2011*
```{r}
b.CI2.imp=
  bugs.imp %>% 
  filter(Code=='CI2')

b.CI2.imp=# add extra row with missing year and NA for spp abundances
b.CI2.imp %>% 
  add_row(Code='CI2', year=1991)

b.CI2.imp=# add extra row with missing year and NA for spp abundances
b.CI2.imp %>% 
  add_row(Code='CI2', year=2009)

b.CI2.imp=# add extra row with missing year and NA for spp abundances
b.CI2.imp %>% 
  add_row(Code='CI2', year=2011)


# sort to order by year
b.CI2.imp=
b.CI2.imp[order(b.CI2.imp$year),]


# impute species as median abundance for the entire time series
b.CI2.imp=
  cbind.data.frame(b.CI2.imp[,c(1,2)],
   apply(b.CI2.imp[,-c(1,2)], 2, NA2ma)
  )

plot(
rowSums(decostand(b.CI2.imp[,-c(1,2)], 'pa')), type='l')


```

#CI4 imputation
*site is missing only 1991*
```{r}
b.CI4.imp=
  bugs.imp %>% 
  filter(Code=='CI4')

b.CI4.imp=# add extra row with missing year and NA for spp abundances
b.CI4.imp %>% 
  add_row(Code='CI4', year=1991)

# sort to order by year
b.CI4.imp=
b.CI4.imp[order(b.CI4.imp$year),]


# impute species as median abundance for the entire time series
b.CI4.imp=
  cbind.data.frame(b.CI4.imp[,c(1,2)],
   apply(b.CI4.imp[,-c(1,2)], 2, NA2ma)
  )

plot(
rowSums(decostand(b.CI4.imp[,-c(1,2)], 'pa')), type='l')


```

#CI5 imputation
*site is missing only 1991, 2009, 2011*
```{r}
b.CI5.imp=
  bugs.imp %>% 
  filter(Code=='CI5')

b.CI5.imp=# add extra row with missing year and NA for spp abundances
b.CI5.imp %>% 
  add_row(Code='CI5', year=1991)


b.CI5.imp=# add extra row with missing year and NA for spp abundances
b.CI5.imp %>% 
  add_row(Code='CI5', year=2009)

b.CI5.imp=# add extra row with missing year and NA for spp abundances
b.CI5.imp %>% 
  add_row(Code='CI5', year=2011)


# sort to order by year
b.CI5.imp=
b.CI5.imp[order(b.CI5.imp$year),]


# impute species as median abundance for the entire time series
b.CI5.imp=
  cbind.data.frame(b.CI5.imp[,c(1,2)],
   apply(b.CI5.imp[,-c(1,2)], 2, NA2ma)
  )

plot(b.CI5.imp$year,
rowSums(decostand(b.CI5.imp[,-c(1,2)], 'pa')), type='l')


```



# Impute LI1 *only 1991 missing*

```{r}
b.LI1.imp=
  bugs.imp %>% 
  filter(Code=='LI1')

b.LI1.imp=# add extra row with missing year and NA for spp abundances
b.LI1.imp %>% 
  add_row(Code='LI1', year=1991)

# sort to order by year
b.LI1.imp=
b.LI1.imp[order(b.LI1.imp$year),]

# impute species as median abundance for the entire time series
b.LI1.imp=
  cbind.data.frame(b.LI1.imp[,c(1,2)],
   apply(b.LI1.imp[,-c(1,2)], 2, NA2ma)
  )

plot(b.LI1.imp$year,
rowSums(decostand(b.LI1.imp[,-c(1,2)], 'pa')), type='l')

```


# Impute LI2 *only 1991 missing*

```{r}
b.LI2.imp=
  bugs.imp %>% 
  filter(Code=='LI2')

b.LI2.imp=# add extra row with missing year and NA for spp abundances
b.LI2.imp %>% 
  add_row(Code='LI2', year=1991)

# sort to order by year
b.LI2.imp=
b.LI2.imp[order(b.LI2.imp$year),]

# impute species as median abundance for the entire time series
b.LI2.imp=
  cbind.data.frame(b.LI2.imp[,c(1,2)],
   apply(b.LI2.imp[,-c(1,2)], 2, NA2ma)
  )

plot(b.LI2.imp$year,
rowSums(decostand(b.LI2.imp[,-c(1,2)], 'pa')), type='l')

```


# Impute LI4 *missing 1991, 2008, 2009, 20011*

```{r}
b.LI4.imp=
  bugs.imp %>% 
  filter(Code=='LI4')

b.LI4.imp=# add extra row with missing year and NA for spp abundances
b.LI4.imp %>% 
  add_row(Code='LI4', year=1991)

b.LI4.imp=# add extra row with missing year and NA for spp abundances
b.LI4.imp %>% 
  add_row(Code='LI4', year=2008)

b.LI4.imp=# add extra row with missing year and NA for spp abundances
b.LI4.imp %>% 
  add_row(Code='LI4', year=2009)

b.LI4.imp=# add extra row with missing year and NA for spp abundances
b.LI4.imp %>% 
  add_row(Code='LI4', year=2011)


# sort to order by year
b.LI4.imp=
b.LI4.imp[order(b.LI4.imp$year),]

# impute species as median abundance for the entire time series
b.LI4.imp=
  cbind.data.frame(b.LI4.imp[,c(1,2)],
   apply(b.LI4.imp[,-c(1,2)], 2, NA2ma)
  )

plot(b.LI4.imp$year,
rowSums(decostand(b.LI4.imp[,-c(1,2)], 'pa')), type='l')

```



# Impute LI6 *missing 1991, 1994*

```{r}
b.LI6.imp=
  bugs.imp %>% 
  filter(Code=='LI6')

b.LI6.imp=# add extra row with missing year and NA for spp abundances
b.LI6.imp %>% 
  add_row(Code='LI6', year=1991)

b.LI6.imp=# add extra row with missing year and NA for spp abundances
b.LI6.imp %>% 
  add_row(Code='LI6', year=1994)


# sort to order by year
b.LI6.imp=
b.LI6.imp[order(b.LI6.imp$year),]

# impute species as median abundance for the entire time series
b.LI6.imp=
  cbind.data.frame(b.LI6.imp[,c(1,2)],
   apply(b.LI6.imp[,-c(1,2)], 2, NA2ma)
  )

plot(b.LI6.imp$year,
rowSums(decostand(b.LI6.imp[,-c(1,2)], 'pa')), type='l')

```


# Impute LI7 *missing 1991, 1994*

```{r}
b.LI7.imp=
  bugs.imp %>% 
  filter(Code=='LI7')

b.LI7.imp=# add extra row with missing year and NA for spp abundances
b.LI7.imp %>% 
  add_row(Code='LI7', year=1991)

b.LI7.imp=# add extra row with missing year and NA for spp abundances
b.LI7.imp %>% 
  add_row(Code='LI7', year=1994)


# sort to order by year
b.LI7.imp=
b.LI7.imp[order(b.LI7.imp$year),]

# impute species as median abundance for the entire time series
b.LI7.imp=
  cbind.data.frame(b.LI7.imp[,c(1,2)],
   apply(b.LI7.imp[,-c(1,2)], 2, NA2ma)
  )

plot(b.LI7.imp$year,
rowSums(decostand(b.LI7.imp[,-c(1,2)], 'pa')), type='l')

```



# Impute LI8 *missing 1991, 1994, 2011*

```{r}
b.LI8.imp=
  bugs.imp %>% 
  filter(Code=='LI8')

b.LI8.imp=# add extra row with missing year and NA for spp abundances
b.LI8.imp %>% 
  add_row(Code='LI8', year=1991)

b.LI8.imp=# add extra row with missing year and NA for spp abundances
b.LI8.imp %>% 
  add_row(Code='LI8', year=1994)


b.LI8.imp=# add extra row with missing year and NA for spp abundances
b.LI8.imp %>% 
  add_row(Code='LI8', year=2011)



# sort to order by year
b.LI8.imp=
b.LI8.imp[order(b.LI8.imp$year),]

# impute species as median abundance for the entire time series
b.LI8.imp=
  cbind.data.frame(b.LI8.imp[,c(1,2)],
   apply(b.LI8.imp[,-c(1,2)], 2, NA2ma)
  )

plot(b.LI8.imp$year,
rowSums(decostand(b.LI8.imp[,-c(1,2)], 'pa')), type='l')

```


## RBIND all imputed sites into one combined df
*this LB10.imp, includes spp observed in >5 streams & 7years, to allow analysis of synchrony using wavelet methods*

```{r}
LB10.imp=rbind.data.frame(b.CI1.imp, b.CI2.imp, b.CI4.imp, b.CI5.imp, b.LI1.imp, b.LI2.imp, b.LI4.imp,
                          b.LI6.imp, b.LI7.imp, b.LI8.imp)

rownames(LB10.imp)=paste(LB10.imp$Code, LB10.imp$year, sep="_")
```


# *How many years for each species in the LB10.imp (some spp are present in <7y)*
I think this is because the vector with years for each spp comes from the complete data (all streams)
So when I use it to subset spp in the imputed df, two species then had lower n of years than that.
```{r}
LB10.imp %>% 
  select(-Code) %>% 
  group_by(year) %>% summarise_all(sum) %>% 
  select(-year) %>% 
  decostand('pa') %>% 
  colSums() %>% 
  sort(decreasing = T)
```





# Create a vector with species identity for the reduced df with 10 streams and no holes
```{r}
speciesID.imp=
colnames(LB10.imp[,-c(1,2)])
```

# The list of *species synchrony imputed* for each species between each stream in the reduced set of sites (LB10.imp). Based on simple correlation
```{r}
spp.list.corr.imp<-list()

for(spp in 1:length(speciesID.imp)) {
  spp.list.corr.imp[[spp]]= 
  #speciesID.subs=as.factor(names(spp_site_occur[spp_site_occur>6]))
  
  sppcorr=
  LB10.imp %>% 
  select(Code, year, speciesID.imp[spp]) %>%  
  spread(Code, value=speciesID.imp[spp]) %>% select(-year) %>% 
  cor(method='spearman') %>% 
  gdata::unmatrix() %>% 
  data.frame() %>% 
  cbind(rep(speciesID.imp[spp], 1 )) %>% 
    cbind(row.names(.)) %>% 
  `colnames<-`(c('synch.imp', 'Species','site.pair'))# use this function: `colnames<-` to assign column names within a pipe
    }
  

spp.list.corr.imp[[12]]

```

# Derive the long format of *species synchrony with imputed values*
*this is like the synchronies_spp but with imputed values from missing years-spp*
```{r}
synchronies_spp.imp=
rbindlist(spp.list.corr.imp)

synchronies_spp.imp=
synchronies_spp.imp %>% 
  filter(synch.imp != 1)

#add column with site-pairs and species
synchronies_spp.imp$site.pair_species=
  paste(synchronies_spp.imp$site.pair, synchronies_spp.imp$Species, sep='_')
```

# add the unique site-pair and species combination for the synchronies_spp data (not the imputed)
```{r}
synchronies_spp$site.pair_species=
  paste(synchronies_spp$site.pair, synchronies_spp$Species, sep='_')
```

# add the imputed species level synchronies to the synchronies_df (where there are info on dist, zonation trt)
```{r}
synchronies_spp$synch.imputed=
  synchronies_spp.imp$synch.imp[match(synchronies_spp$site.pair_species, synchronies_spp.imp$site.pair_species)]
```

# Nice correlation between *imputed and observed synchrony for each species*
```{r}
synchronies_spp %>% 
  ggplot()+aes(synch, synch.imputed)+geom_point()

summary(lm(synch~synch.imputed, data=synchronies_spp))

```


# Create a list similar to spp.list.corr but as simple correlation matrix. Used for calculating the mean across species.

```{r}
list2.imp<-list()

for(spp in 1:length(speciesID.imp)) {
  list2.imp[[spp]]= 
  #speciesID.subs=as.factor(names(spp_site_occur[spp_site_occur>6]))
  
  sppcorr=
  LB10.imp %>% 
  select(Code, year, speciesID.imp[spp]) %>%  
  spread(Code, value=speciesID.imp[spp]) %>% select(-year) %>% 
  cor(method='spearman')
}

list2.imp[41]
```

# Derive the mean synchrony across sites based on the mean synchrony of species in LB10 imputed missing years
```{r}
array2.imp=array(as.numeric(unlist(list2.imp)), dim=c(length(unique(LB10.imp$Code)), length(unique(LB10.imp$Code)), length(speciesID.imp)))
dimnames(array2.imp)=list( unique(LB10.imp$Code), unique(LB10.imp$Code),speciesID.imp )

array2.imp[,,43]

```

# The mean synchrony across LB10 imputed sites based on simple correlation across populations from species occurring > 6 streams
```{r}
mean_spp_corr.mat.imp=
apply(array2.imp, c(1,2), mean, na.rm=T)
```

# *Plotting the man pop synchrony with imputated years shows no effect of imputation, if anything, overall synch values seems lower!*
```{r fig.width=7}
#library(corrplot)
corrplot(round(mean_spp_corr.mat.imp,2), method='color', type='lower', addCoef.col = "black", title = 'Mean pop synchrony (corr-imputed)', mar=c(0,0,2,0), number.cex = .8, tl.cex = 0.8)
```



# extract diversity specific timeseries *needed to claculate on loops etc*
*abund.ts*
```{r}
richness.ts=
lb10.div %>% 
  select(c(Code,year,richness)) %>%
  pivot_wider(names_from = year, values_from = richness) %>% as.data.frame()

shannon.ts=
lb10.div %>% 
  select(c(Code,year,shannon)) %>%
  pivot_wider(names_from = year, values_from = shannon) %>% as.data.frame()


simpson.ts=
lb10.div %>% 
  select(c(Code,year,simpson)) %>%
  pivot_wider(names_from = year, values_from = simpson) %>% as.data.frame()

rownames(richness.ts)=richness.ts$Code
rownames(shannon.ts)=shannon.ts$Code
rownames(simpson.ts)=simpson.ts$Code

richness.ts$Code=NULL
shannon.ts$Code=NULL
simpson.ts$Code=NULL


abund.ts=
  lb10.div %>% 
  select(c(Code, year, abund)) %>% 
  pivot_wider(names_from = year, values_from = abund) %>% as.data.frame()

abund.ts$Code=NULL

```


###############################
## The moving window synchrony
###############################

#function that calculates the population synch for each spp and rbindlist the list
```{r}
pop_synch_f=function(df){
  testlist<-list()

for(spp in 1:length(speciesID.subs)) {
testlist[[spp]]= 
  #speciesID.subs=as.factor(names(spp_site_occur[spp_site_occur>6]))
  
  #sppcorr=
  df %>% 
  select(Code, year, speciesID.subs[spp]) %>%  
  spread(Code, value=speciesID.subs[spp]) %>% select(-year) %>% 
  cor(method='spearman') %>% 
  gdata::unmatrix() %>% 
  data.frame() %>% 
  cbind(rep(speciesID.subs[spp], 1 )) %>% 
    cbind(row.names(.)) %>% 
  `colnames<-`(c('synch', 'Species','site.pair'))# use this function: `colnames<-` to assign column names within a pipe
  
 }
 
  res=data.table::rbindlist(testlist) 
  res=res %>%  filter(synch != 1)
  res$geodist=dist.pairs$dist[match(res$site.pair, dist.pairs$site_pairs)] # add the geodist
  return(res)
  
}
```


# create a timewindow list to include year given year in the window, needed to subset df while running the synchrony function
```{r}
time.windowlist=list()
ywindow=5
for(i in 1:(length(yearID)-ywindow)){
time.windowlist[[i]]=  (yearID[i:(i+5)])}

```

# run the pop_synch function over each yeargroup in window ,as defined by time.windolist
```{r}

pop_synchronies_mw.list=list()

for(mwy in 1:length(time.windowlist)){
pop_synchronies_mw.list[[mwy]]=
pop_synch_f(LB10.imp %>% filter(year %in% time.windowlist[[mwy]]))
}

pop_synchronies_mw.list[[24]]

```

# Now calculate the man across species of synchrony for each pairwise site
```{r}
meanpop_synchronies_mw.list=list()

for(mwy in 1:length(time.window5)){
meanpop_synchronies_mw.list[[mwy]]=
pop_synchronies_mw.list[[mwy]] %>% 
  group_by(site.pair) %>% 
  summarise(mean.pop.synch=mean(synch), geodist=mean(geodist)) %>% 
  mutate(timewindow=rep(time.window5[[mwy]], 90)) # add timewindow info 90 rows for each yearwindow (site-pairs)
         }

meanpop_synchronies_mw.list[[1]]

```

# Rbind list of the moving window population synchr and add timewindow and NAO values
```{r}
meanpop_synchronies_mw=
rbindlist(meanpop_synchronies_mw.list) %>% 
  mutate(timewindow=as.factor(timewindow)) 

meanpop_synchronies_mw$NAOmw=wNAO_window$NAOwindow[match(meanpop_synchronies_mw$timewindow, wNAO_window$time.window5)]

```

# establish quantile values for NAOmw
```{r}
library(gtools)
#meanpop_synchronies_mw$NAOmw.q4=quantcut(meanpop_synchronies_mw$NAOmw, q=4)

#meanpop_synchronies_mw$NAOmw.q6=quantcut(meanpop_synchronies_mw$NAOmw, q=6)

meanpop_synchronies_mw$NAOmw.q5=quantcut(meanpop_synchronies_mw$NAOmw, q=5)




```

# Plot of mean population synchorny moving window vs quantile NAO
```{r fig.width=5.5}
meanpop_synchronies_mw %>% 
ggplot()+aes(geodist, mean.pop.synch, col=NAOmw.q4)+geom_smooth(method='lm', se=F)+scale_color_brewer (palette='Reds')+
  ylab('Mean population synchrony (mw)')+xlab('Distance')+geom_smooth(col='grey40', method='lm', size=2, linetype='dashed')

meanpop_synchronies_mw %>% 
ggplot()+aes(geodist, mean.pop.synch, col=NAOmw.q4)+geom_smooth(se=F)+scale_color_brewer (palette='Reds')+
  ylab('Mean population synchrony (mw)')+xlab('Distance')


meanpop_synchronies_mw %>% 
ggplot()+aes(geodist, mean.pop.synch, col=NAOmw.q5)+geom_smooth(method='lm', se=F)+scale_color_brewer (palette='Reds')+
  ylab('Mean population synchrony (mw)')+xlab('Distance')+geom_smooth(col='grey40', method='lm', size=1.5, linetype='dashed')+
  labs(col='NAO quantiles')


```

# *Plot for Fig.3 using all timewindows
*not in paper and not reflecting the model*
```{r}
#pdf("Fig.3_Synch_slopes.pdf", w=5.5, h=4)
meanpop_synchronies_mw %>% 
ggplot()+aes(geodist, mean.pop.synch, col=NAOmw.q5)+geom_smooth(method='lm', se=F)+scale_color_brewer (palette='Reds')+
  ylab('Mean population synchrony (mw)')+xlab('Distance')+geom_smooth(col='grey40', method='lm', size=1.5, linetype='dashed')+
  labs(col='NAO quantiles')

#dev.off()
```




# simple model of synchrony ~ geodistance*NAO interaction
positive interaction effect but weak when including every second window, and removing duplicated pairs*
*not in paper*
```{r}
meanpop_synchronies_mw

m0=glm(mean.pop.synch~geodist*NAOmw, data=meanpop_synchronies_mw, family='gaussian')
summary(m0)

m.int=glm(mean.pop.synch~geodist*NAOmw, 
       data=meanpop_synchronies_mw[!duplicated(meanpop_synchronies_mw$mean.pop.synch),] %>% 
         filter(timewindow %in% time.window5[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29)]),
       family = 'gaussian')

summary(m.int)

```



# Model random effect of NAO quantile to mirror the plot with decay...
*Support for the inclusion of random INTERCEPT (less so for random slope) effect for different NAO quantiles* 
```{r}
library(nlme)

# no random eff at all
m0= gls(mean.pop.synch ~ geodist,
        data= meanpop_synchronies_mw[!duplicated(meanpop_synchronies_mw$mean.pop.synch),] %>% 
          filter(timewindow %in% time.window5[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29)]) , method='REML')
summary(m0)

# random intercept for NAO quantiles
m1=
  lme(mean.pop.synch ~ geodist , random=~1|NAOmw.q5, 
      data= meanpop_synchronies_mw[!duplicated(meanpop_synchronies_mw$mean.pop.synch),] %>% 
          filter(timewindow %in% time.window5[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29)]) )

#random intercept and slope
m2=
  lme(mean.pop.synch ~ geodist , random=~1+geodist|NAOmw.q5, 
      data= meanpop_synchronies_mw[!duplicated(meanpop_synchronies_mw$mean.pop.synch),] %>% 
         filter(timewindow %in% time.window5[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29)]))
        


summary(m1)
summary(m2)

# support for inclusion of random intercept
anova(m0,m1,m2)



```





# create subset clean (model like) version of mean population synchrony mw, for plotting model fits more easily
Removing duplicated synchrony values and skipping every second time window (as done in the models)
```{r}
meanpop_synchronies_mw_model= meanpop_synchronies_mw[!duplicated(meanpop_synchronies_mw$mean.pop.synch),] %>% 
         filter(timewindow %in% time.window5[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29)])

meanpop_synchronies_mw_model$m1fit<-
predict(m1)

meanpop_synchronies_mw_model$m2fit<-
predict(m2)


```

# plotting *Fig4 with equal slope* for NAO quantiles (ex fig.3)
```{r fig.width=5.5}
pdf("Fig.4_intercept_revis.pdf", w=5.5, h=4)
meanpop_synchronies_mw_model %>% 
ggplot()+aes(geodist, mean.pop.synch, col=NAOmw.q5)+geom_line(aes(y=m1fit), size=1.2)+scale_color_brewer (palette='Reds')+
  ylab('Mean population synchrony (mw)')+xlab('Distance')+geom_smooth(col='grey10', method='lm', size=1, linetype='dashed')+
  labs(col='NAO quantiles')
dev.off()
```


#model m2 (random slope) Plotting
```{r}
meanpop_synchronies_mw_model %>% 
ggplot()+aes(geodist, mean.pop.synch, col=NAOmw.q5)+geom_line(aes(y=m2fit), size=1.2)+scale_color_brewer (palette='Reds')+
  ylab('Mean population synchrony (mw)')+xlab('Distance')+geom_smooth(col='grey10', method='lm', size=1, linetype='dashed')+
  labs(col='NAO quantiles')
```





#########################################
# Calculate Abundance synchrony moving window
#########################################

# abundance synchrony mw
```{r}
abund.synch_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=1))


rownames(abund.synch_mw)=time.window5
colnames(abund.synch_mw)='Abund.synch.mw'



# the abund synch as simple cor
for(y in 1: length(time.window5)){
 abund.synch_mw[y,]= mean(
   cor(
     t(abund.ts[,c(y:(y+5))]), method='spearman'
     )
   )
}
```




add moving window nao
```{r}
abund.synch_mw$wNAOwindow=wNAO_window$NAOwindow[match(rownames(abund.synch_mw), wNAO_window$time.window5)]
```

# calculate the variance of the abundance synchronmy mw
```{r}
abund.synch.var_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=1))


rownames(abund.synch.var_mw)=time.window5
colnames(abund.synch.var_mw)='Abund.synch.var.mw'



# the abund synch as simple cor
for(y in 1: length(time.window5)){
 abund.synch.var_mw[y,]= var(as.vector(
   cor(
     t(abund.ts[,c(y:(y+5))]), method='spearman'
     )
   ))
}
```
# add variance in abund synchrony
```{r}
abund.synch_mw$synch.var=abund.synch.var_mw$Abund.synch.var.mw
```


# Plot abundance synchrony mw vs NAO

```{r fig.width=6}


abund.synch_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>% 
  #mutate(RegRich.mw= reg.richness %>% rollmean(k=6, align = 'center')) %>% 
  #mutate(MeanRich.mw= rowMeans(Richness.mw)) %>% 
  ggplot()+aes(wNAOwindow, Abund.synch.mw, label=row.names(abund.synch_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),]))+geom_point(size=2)+geom_text_repel(size=3)+
  geom_smooth( col='grey50', alpha=.3, method='lm')+ylab('Abundance synchrony mw')+ xlab('Winter NAO mw')


```



# Plot the abundance mw over time along with NAO window
```{r fig.width=6}

abund.synch_mw %>% 
  ggplot()+aes(row.names(abund.synch_mw), Abund.synch.mw, group=1)+geom_point(size=3)+theme_bw()+
  geom_path(aes(row.names(abund.synch_mw) ,wNAOwindow), col='blue')+theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab(NULL)+ylab('Mean abundance snchrony')+scale_y_continuous(sec.axis = sec_axis(~.*1, name = "Winter NAO mw"))+theme(axis.title.y.right = element_text(color = "blue")) +theme(axis.text.y.right = element_text(color='blue'))+theme(axis.title = element_text(size=12), axis.text.x = element_text(size=12)) +scale_x_discrete(labels=time.window5_skip)

abund.synch_mw %>% 
  ggplot()+aes(row.names(abund.synch_mw), Abund.synch.mw, group=1)+geom_point(size=3)+theme_bw()+geom_errorbar(aes(ymin=Abund.synch.mw-synch.var, ymax=Abund.synch.mw+synch.var), col='grey60')+
  geom_path(aes(row.names(abund.synch_mw) ,wNAOwindow), col='blue')+theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab(NULL)+ylab('Mean abundance snchrony')+scale_y_continuous(sec.axis = sec_axis(~.*1, name = "Winter NAO mw"))+theme(axis.title.y.right = element_text(color = "blue")) +theme(axis.text.y.right = element_text(color='blue'))+theme(axis.title = element_text(size=12), axis.text.x = element_text(size=12)) +scale_x_discrete(labels=time.window5_skip)


```

# Save plot abundance synchrony
# Plot *part of Fig.5 (ex fig.4)* 
```{r}
pdf('Abund_synch_time.pdf', w=6, h=3.5)
abund.synch_mw %>% 
  ggplot()+aes(row.names(abund.synch_mw), Abund.synch.mw, group=1)+geom_point(size=3)+theme_bw()+geom_errorbar(aes(ymin=Abund.synch.mw-synch.var, ymax=Abund.synch.mw+synch.var), col='grey60', size=0.6)+
  geom_point(size=3)+
  geom_path(aes(row.names(abund.synch_mw) ,wNAOwindow), col='blue')+theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab(NULL)+ylab('Mean abundance snchrony')+scale_y_continuous(sec.axis = sec_axis(~.*1, name = "Winter NAO mw"))+theme(axis.title.y.right = element_text(color = "blue")) +theme(axis.text.y.right = element_text(color='blue'))+theme(axis.title = element_text(size=12), axis.text.x = element_text(size=12)) +scale_x_discrete(labels=time.window5_skip)

dev.off()
```



# Modelling abundance synchrony vs NAO
```{r}
#library(nlme)
# the proper model for abundance
m_abun=gls(Abund.synch.mw~wNAOwindow,data=abund.synch_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),], cor=corAR1(0.7))
summary(m_abun)  # ab.synch = 0.26 +0.35 * NAOw
```

# Nao phases not related to abundance absolute values, but just its spatial synchr across streams
```{r}

lb10.div %>% 
  ggplot()+aes(NAOw, log(abund), col=Code)+geom_smooth(method='lm', se=F)+geom_point()
```




#############################
### Wavelet linear model ####
# *testing effect of NAO vs abundance*

```{r}
abund.ts
t_NAOw_85.18.exp

dat=list(as.matrix(abund.ts), t_NAOw_85.18.exp)
dat=lapply(FUN=function(x){cleandat(x,time.lb10,1)$cdat},X=dat)

```


```{r}
wlm_nao_abu=wlm(dat, time.lb10, resp=1, pred=2, norm='powall' )

wlm_nao_abu$coefs
summary(wlm_nao_abu)
wlm_nao_abu$coher

syncexpl(wlm_nao_abu)
wlm_nao_abu$wts


```


# *The predicted synchrony in abundance from NAO alone*
```{r}
predabund_sync=predsync(wlm_nao_abu)

plotmag(predabund_sync, title='NAO-Predicted synchrony abundance')

plot_wlm_predict=plotmag(predabund_sync, title='NAO-Predicted synchrony abundance')
  

```

```{r}

plotmag(
    wpmf(
cleandat(
lb10.div %>% 
  select(c(Code,year,abund)) %>%
  pivot_wider(names_from = year, values_from = abund) %>% select(-Code) %>% as.matrix(),
   times=time.lb10, clev=4)$cdat, times=time.lb10, sigmethod = 'quick'), title='Wpmf Abundance')



```

# export wavelet model prediction plot
#Plot *part of Fig.5 (wavelet model)*
```{r}
pdf('plot_wlm_predict.pdf', w=6, h=4.5)
plotmag(predabund_sync, title='NAO-Predicted synchrony abundance')
dev.off()
```

# export wavelet model prediction plot
#Plot *part of Fig.5 (wawelet observed synch)*
```{r}
pdf('plot_wpmf_abund.pdf', w=6, h=4.5)

plotmag(
    wpmf(
cleandat(
lb10.div %>% 
  select(c(Code,year,abund)) %>%
  pivot_wider(names_from = year, values_from = abund) %>% select(-Code) %>% as.matrix(),
   times=time.lb10, clev=4)$cdat, times=time.lb10, sigmethod = 'quick'), title='Wpmf Abundance')

dev.off()
```


##
```{r}
#library(wsyn)

bandtest(wlm_nao_abu, c(4,6))

```


# Extract the yearly temperature series, to use in wavelet modelling along NAO
```{r}
Tyearly.wide=
Tyearly %>% 
  pivot_wider(values_from = mean.w.T, names_from = Site) %>% as.data.frame()

rownames(Tyearly.wide)=Tyearly.wide$Ante
Tyearly.wide$Ante=NULL
Tyearly.wide

```
# that is the trasposed df to use
```{r}
t(Tyearly.wide)
```

```{r}
library(wsyn)
```

# Prepare for linear wavelet model using also Temperature 
```{r}
dat2=list(as.matrix(abund.ts), t_NAOw_85.18.exp, t(Tyearly.wide))
dat2=lapply(FUN=function(x){cleandat(x,time.lb10,1)$cdat},X=dat2)
```

# Wavelet model including Tempe does not add much explained variation. 
```{r}
wlm_nao.Temp_abu=wlm(dat2, time.lb10, resp=1, pred=c(2,3), norm='powall' )


summary(wlm_nao.Temp_abu)
wlm_nao_abu$coher

syncexpl(wlm_nao.Temp_abu)
```

```{r}
plotmag(predsync(wlm_nao.Temp_abu))
```


