---
title: "GCB_StabilityPartition"
author: "Stefano Larsen"
date: "2023-08-16"
output: html_document
---



```{r}

install.packages('ggrepel')
library(readxl)
library(tidyverse)
library(reshape2)
library(ggrepel)
library(zoo)
library(vegan)
library(codyn)
library(iCAMP)
library(data.table)
install.packages('gdata')
library(gdata)
library(corrplot)
library(RColorBrewer)
```



```{r}
theme_set(theme_bw())
```


#######################
## Hieratchical stability partitioning (Wang's style)
######################


# This is the original *Wang's code for partitioning*
```{r}
var.partition <-function(metacomm_tsdata){
## The function "var.partition" performs the partitioning of variability 
## across hierarchical levesl within a metacommunity.
## The input array "metacomm_tsdata" is an N*T*M array. The first dimension represents N species, 
## the second represents time-series observations of length T, and the third represents M local communities. 
## The output includes four variability and four synchrony metrics as defined in the main text.
## Note that, to be able to handle large metacommunities, this code has avoided calculating all covariance.

ts_metacom <-apply(metacomm_tsdata,2,sum,na.rm=T)
ts_patch <-apply(metacomm_tsdata,c(2,3),sum,na.rm=T)
ts_species <-apply(metacomm_tsdata,c(1,2),sum,na.rm=T)
sd_metacom <-sd(ts_metacom,na.rm=T)
sd_patch_k <-apply(ts_patch,2,sd,na.rm=T)
sd_species_i <-apply(ts_species,1,sd,na.rm=T)
sd_species_patch_ik <-apply(metacomm_tsdata,c(1,3),sd,na.rm=T)
mean_metacom <-mean(ts_metacom,na.rm=T)

CV_S_L <-sum(sd_species_patch_ik,na.rm=T)/mean_metacom  #local-scale average species variability (weighted average of local                                                             population variability across sites and species)
CV_C_L <-sum(sd_patch_k,na.rm=T)/mean_metacom            #local-scale average community variability (weighted average of community                                                             variability across sites)
CV_S_R <-sum(sd_species_i,na.rm=T)/mean_metacom          #regional-scale average species variability (weighted average of                                                                       metapopulation variability across species)
CV_C_R <-sd_metacom/mean_metacom                  #metacommunity variability (CV total metacommunity biomass)

phi_S_L2R <-CV_S_R/CV_S_L  #species-level spatial synchrony (weighted average intraspecific across species)
phi_C_L2R <-CV_C_R/CV_C_L  #community-level spatial synchrony  (of total community biomass across sites)
phi_S2C_L <-CV_C_L/CV_S_L  #local-scale species synchrony     (weighted average interspecific across patches)
phi_S2C_R <-CV_C_R/CV_S_R   #regional-scale species synchrony  (among different meta-populations [sum of covariances                                                                                                           beween species]
partition_3level <-c(CV_S_L=CV_S_L, CV_C_L=CV_C_L, CV_S_R=CV_S_R, CV_C_R=CV_C_R, phi_S_L2R=phi_S_L2R, phi_C_L2R=phi_C_L2R, phi_S2C_L=phi_S2C_L, phi_S2C_R=phi_S2C_R)
return(partition_3level)
}
```


## Convert the bio data into an array to feed the var.partition function. Will use the imputed data this time on LB10



# gather bugs data in long format -useful for subsequent manipulations
```{r}
bugs.long.imp=
LB10.imp %>% 
  gather(Species, abund, -c(Code, year))
```

# code, year and species to factor
```{r}
bugs.long.imp$Code=as.factor(bugs.long.imp$Code)
bugs.long.imp$year=as.factor(bugs.long.imp$year)
bugs.long.imp$Species=as.factor(bugs.long.imp$Species)
```

# define the number of years, sites and species overall
```{r}
nlevels(bugs.long.imp$year)
nlevels(bugs.long.imp$Code)
nlevels(bugs.long.imp$Species)
```

# create vector keeping year, site and species ID - useful for working in loops
```{r}
yearID=sort(unique(bugs.long.imp$year))
codeID=sort(unique(bugs.long.imp$Code))
speciesID=sort(unique(bugs.long.imp$Species))
```

# Create a list keeping matrix of species x year - each spp a timeseries for each stream
```{r}
mylist=list()
```

```{r}

for(code in 1:nlevels(bugs.long.imp$Code)){
  yearID=sort(unique(bugs.long.imp$year))
  codeID=sort(unique(bugs.long.imp$Code))
  speciesID=sort(unique(bugs.long.imp$Species))
  bugs.long.code=bugs.long.imp[bugs.long.imp$Code==codeID[code],]
  bugs.long.mat.code=(spread(bugs.long.code[,-1], year, abund, drop=F, fill=0))
   rownames(bugs.long.mat.code)=bugs.long.mat.code[,1]
   bugs.long.mat.code[,1]=NULL
 mylist[[code]]=bugs.long.mat.code
}
```

#  see the list
```{r}
mylist[[2]]
```

# Convert the list of species x year (for each stream) into an *ARRAY* with dim = species x years x site
*this could be useful for the CV stability function of Wang*
*note that all spp are included here 65, not only 60*
```{r}
myarray=array(as.numeric(unlist(mylist)), dim=c(nlevels(bugs.long.imp$Species), nlevels(bugs.long.imp$year), nlevels(bugs.long.imp$Code)))
dimnames(myarray)=list(speciesID, yearID, codeID)

myarray[,,1]
myarray[,,3]

dim(myarray)# 60 species, 34 years, 10 sites (with no holes, i.e. imputed)
```

# can use cor between two sites df, to get the cerrelation (synch) between each spp (not only intra-spp cor)
```{r}
# spp X years X sites
dim(myarray)
# here the correlation between site 1 and 2, for each spp pair (including same spp)
xx=cor(t(myarray[,,1]), t(myarray[,,2]))

```


## *Nest the lb10.imp over each site*
```{r}
LB10.imp.n=LB10.imp %>% 
  group_by(Code) %>% 
  nest()
```


# Working with the cody::stability and synchrony functions
# *Stability function* to run over moving window of n years
```{r}

# this function only works within nested version of lb10 nested for sites

stabf=function(df){
df %>% 
  #filter(Code==lb10.siteID[1]) %>% 
  pivot_longer(col=speciesID.imp) %>% # remove also 
  #select(-Code) %>% 
  mutate(name=as.factor(name), year=as.integer(year)) %>% 
 community_stability (time.var = 'year',  abundance.var = 'value')
}


stabf(b.CI1.imp)
```

#tests
```{r}

b.CI1.imp %>% 
  pivot_longer(col=speciesID.imp)

LB10.imp.n$data[[1]] %>% 
  pivot_longer(col=speciesID.imp)

# the stability funciton b works on both the nested and original site data
stabf(b.CI1.imp) # 

stabf(LB10.imp.n$data[[1]])

```



# run the stability function on each nested site
```{r}
LB10.imp.n=
LB10.imp.n %>% 
  mutate(com.stability=unlist(map(data, stabf)))


LB10.imp.n$data[[1]]
```



# function that claculate the comm stability metric *for a window of 6 years (moving by 1y each time)*


```{r}



stabf.mw=function(df){
zio=list()
ywindow=5
for(i in 1:(nrow(df)-ywindow)){
zio[[i]]=  stabf(df[i:(i+5),])}
return(unlist(zio))
}

stabf.mw(b.CI1.imp)

stabf.mw(LB10.imp.n$data[[1]])

```

```{r}
stabf.mw(b.LI7.imp)
```




# run the moving window stability function on each site (nested)
```{r}

LB10.imp.n =
LB10.imp.n %>% 
  mutate(stability.mw=(map(data, stabf.mw)))

LB10.imp.n$stability.mw[[1]]

rbindlist(list(LB10.imp.n$stability.mw))

```

#extract the moving window of comm stability into a df
```{r}

time.window5=matrix(nrow=29)
for(i in 1:29){
time.window5[i,]=paste(time.lb10[i], time.lb10[i+5], sep='_')  
}

comm.stab.mw=cbind.data.frame(time.window=time.window5,
  rbindlist(list(LB10.imp.n$stability.mw))
)

colnames(comm.stab.mw)[2:11]=rownames(stability.metrics)
```




# *Synchrony function* to run over moving window of n years
```{r}

# this function only works within nested version of lb10 nested for sites

synchf=function(df){
df %>% 
  #filter(Code==lb10.siteID[1]) %>% 
  pivot_longer(col=speciesID.imp) %>% # remove also 
  #select(-Code) %>% 
  mutate(name=as.factor(name), year=as.integer(year)) %>% 
 synchrony (time.var = 'year',species.var = 'name',  abundance.var = 'value', metric = 'Loreau')
}


synchf(b.CI1.imp)
```

# synchrony funciton over 5 y moving window (as done fo stability)
```{r}

synchf.mw=function(df){
zio=list()
ywindow=5
for(i in 1:(nrow(df)-ywindow)){
zio[[i]]=  synchf(df[i:(i+5),])}
return(unlist(zio))
}
```



# run the moving window *stability function (codyn) on each site (nested)*
```{r}

LB10.imp.n =
LB10.imp.n %>% 
  mutate(synch.loreau.mw=(map(data, synchf.mw)))

LB10.imp.n$synch.loreau.mw[[1]]

rbindlist(list(LB10.imp.n$stability.mw))
```

# extractt he synchrony moving window loreau
```{r}
synch.loreau.mw=cbind.data.frame(time.window=time.window5,
  rbindlist(list(LB10.imp.n$synch.loreau.mw))
)

colnames(synch.loreau.mw)[2:11]=rownames(stability.metrics)
```




# *Synchrony function (codyn)* to run over moving window of n years
```{r}

# this function only works within nested version of lb10 nested for sites

synchf=function(df){
df %>% 
  #filter(Code==lb10.siteID[1]) %>% 
  pivot_longer(col=speciesID.imp) %>% # remove also 
  #select(-Code) %>% 
  mutate(name=as.factor(name), year=as.integer(year)) %>% 
 synchrony (time.var = 'year',species.var = 'name',  abundance.var = 'value', metric = 'Loreau')
}


synchf(b.CI1.imp)
```

# synchrony funciton over 5 y moving window (as done fo stability)
```{r}

synchf.mw=function(df){
zio=list()
ywindow=5
for(i in 1:(nrow(df)-ywindow)){
zio[[i]]=  synchf(df[i:(i+5),])}
return(unlist(zio))
}
```



# run the *moving window synchrony function (codyn) on each site (nested)*
```{r}

LB10.imp.n =
LB10.imp.n %>% 
  mutate(synch.loreau.mw=(map(data, synchf.mw)))

LB10.imp.n$synch.loreau.mw[[1]]

rbindlist(list(LB10.imp.n$stability.mw))
```

# extractt he synchrony moving window loreau
```{r}
synch.loreau.mw=cbind.data.frame(time.window=time.window5,
  rbindlist(list(LB10.imp.n$synch.loreau.mw))
)

colnames(synch.loreau.mw)[2:11]=rownames(stability.metrics)
```




### *Wang decmposition using moving window*

```{r}
wang_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=length(res.wang.part)))
rownames(wang_mw)=time.window5
colnames(wang_mw)=names(res.wang.part)

var.partition(myarray[,c(1: (1+5)),])


for(y in 1: length(time.window5)){
 wang_mw[y,]= var.partition(myarray[,c(y: (y+5)),])
}

```

# add wNAOwindow on the wang's stability window
```{r}
wang_mw$wNAO.window=wNAO_window$NAOwindow
```

# *wNAO affects all components of Wang stability*
Not in paper
```{r}
wang_mw %>% 
  pivot_longer(cols=c(1:8)) %>% 
  ggplot()+aes(wNAO.window, value)+geom_point()+geom_smooth(method='lm')+facet_wrap(~name, scales = 'free')
```

# Plot of NAO effect on Wangs stability across levels
```{r fig.width=5.5, fig.height=4.4}
wang_mw[c(1,3,4,9)] %>% 
  pivot_longer(cols=c(1:3)) %>% 
  mutate(name=factor(name, levels=c('CV_S_L', 'CV_S_R', 'CV_C_R'))) %>% 
  ggplot()+aes(wNAO.window, value, col=name)+geom_point(size=2, alpha=.2)+
  #geom_line(size=1.2, alpha=0.3)+  
  scale_y_log10()+
  
  geom_smooth(method='lm', se=F)+theme(legend.position = 'top')+ylab('Variability (log10)')+theme(axis.title = element_text(size=12))+
  scale_color_manual(values=c("orange", 'forestgreen', 'grey20'), labels=c('CV population', 'CV metapopulation', 'CV metacommunity'), name='CVcomponent')+theme(legend.title = element_blank())+ylab('Variability')+xlab('Winter NAO window')

  



```

# same plot as above with dashed line for non sign regression
```{r fig.width=5.5, fig.height=4.4}
wang_mw[c(1,3,4,9)] %>% 
  pivot_longer(cols=c(1:3)) %>% 
  mutate(name=factor(name, levels=c('CV_S_L', 'CV_S_R', 'CV_C_R'))) %>% 
  ggplot()+aes(wNAO.window, value, col=name, linetype=name)+geom_point(size=2, alpha=.5)+
  #geom_line(size=1.2, alpha=0.3)+  
  scale_y_log10()+
  
  geom_smooth(method='lm', se=F )+theme(legend.position = 'top')+ylab('Variability (log10)')+theme(axis.title = element_text(size=12))+
  scale_color_manual(values=c("orange", 'forestgreen', 'grey20'), labels=c('CV population', 'CV metapopulation', 'CV metacommunity'), name='CV component')+theme(legend.title = element_blank())+ylab('Variability')+xlab('Winter NAO window')+scale_linetype_manual(values=c('dashed', 'solid','solid'), name='CV component')+guides(linetype='none')

```

# Save CV plot 
```{r}
Plot_CVs=
wang_mw[c(1,3,4,9)] %>% 
  pivot_longer(cols=c(1:3)) %>% 
  mutate(name=factor(name, levels=c('CV_S_L', 'CV_S_R', 'CV_C_R'))) %>% 
  ggplot()+aes(wNAO.window, value, col=name, linetype=name)+geom_point(size=2, alpha=.5)+
  #geom_line(size=1.2, alpha=0.3)+  
  scale_y_log10()+
  
  geom_smooth(method='lm', se=F )+theme(legend.position = 'top')+ylab('Variability (log10)')+theme(axis.title = element_text(size=12))+
  scale_color_manual(values=c("orange", 'forestgreen', 'grey20'), labels=c('CV population', 'CV metapopulation', 'CV metacommunity'), name='CV component')+theme(legend.title = element_blank())+ylab('Variability')+xlab('Winter NAO window')+scale_linetype_manual(values=c('dashed', 'solid','solid'), name='CV component')+guides(linetype='none')


```

# Save plot CVs, skipping years
```{r}
Plot_CVs.skip=
wang_mw[ c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),c(1,3,4,9)] %>% 
  pivot_longer(cols=c(1:3)) %>% 
  mutate(name=factor(name, levels=c('CV_S_L', 'CV_S_R', 'CV_C_R'))) %>% 
  ggplot()+aes(wNAO.window, value, col=name, linetype=name)+geom_point(size=2, alpha=.5)+
  #geom_line(size=1.2, alpha=0.3)+  
  scale_y_log10()+
  
  geom_smooth(method='lm', se=F )+theme(legend.position = 'top')+ylab('Variability (log10)')+theme(axis.title = element_text(size=12))+
  scale_color_manual(values=c("orange", 'forestgreen', 'grey20'), labels=c('CV population', 'CV metapopulation', 'CV metacommunity'), name='CV component')+theme(legend.title = element_blank())+ylab('Variability')+xlab('Winter NAO window')+scale_linetype_manual(values=c('dashed', 'solid','solid'), name='CV component')+guides(linetype='none')
```




# Autocorrelation models for variability CV (pop, metapop, metacommunity - Wang) vs NAO
```{r}
mod.CV_S_L=
  gls(CV_S_L~wNAO.window, data=wang_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),], cor=corAR1(0.6))
# population CV barely sig
summary(mod.CV_S_L)
 

mod.CV_S_R=
  gls(CV_S_R~wNAO.window, data=wang_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),], cor=corAR1(0.8))

summary(mod.CV_S_R)

mod_CV_C_R=
gls(CV_C_R~wNAO.window, data=wang_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),], cor=corAR1(0.6))

summary(mod_CV_C_R)

```



# Plot of community synchorny (phi) at local and regional scale
*note here I also add funcitonal diversity althought not still present in the workflow in theory*
*also all data are plotted here, not skipping years*
```{r}
cbind.data.frame(wang_mw, FunRao.mw=rowMeans(FunRao.mw)) %>% 
  pivot_longer(cols=c(7,8), names_to = 'SynchComponent') %>% 
  mutate(SynchComponent=factor(SynchComponent)) %>% 
  mutate(SynchComponent_r=recode_factor(SynchComponent, phi_S2C_L='Local_Synchony', phi_S2C_R='Regional_Synchrony')) %>% 
  ggplot()+aes(wNAO.window, value )+geom_point(size=2)+geom_smooth(method='lm')+
  facet_wrap(~SynchComponent_r)+theme(legend.position = 'right')+ylab('Community synchrony')+theme(axis.title = element_text(size=12))+xlab('Winter NAO window')
```





# Save plot of Synchronies
```{r}
plot_synchronies=
cbind.data.frame(wang_mw, FunRao.mw=rowMeans(FunRao.mw)) %>% 
  pivot_longer(cols=c(7,8), names_to = 'SynchComponent') %>% 
  mutate(SynchComponent=factor(SynchComponent)) %>% 
  mutate(SynchComponent_r=recode_factor(SynchComponent, phi_S2C_L='Local Synchony', phi_S2C_R='Regional Synchrony')) %>% 
  ggplot()+aes(wNAO.window, value )+geom_point(size=2, alpha=0.5)+geom_smooth(method='lm')+
  facet_wrap(~SynchComponent_r)+theme(legend.position = 'right')+ylab('Community synchrony')+theme(axis.title = element_text(size=12))+xlab('Winter NAO window')
```


```{r fig.width=5, fig.height=5}
#library(patchwork)
plot_synchronies /
  Plot_CVs+plot_annotation(tag_levels = 'A')
```



# Save plot synchronies, skipping years
```{r}
plot_synchronies.skip=
 wang_mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),]   %>% 
  select(phi_S2C_L, phi_S2C_R, wNAO.window) %>% 
  pivot_longer(cols=-3, names_to = 'SynchComponent') %>% 
  mutate(SynchComponent=factor(SynchComponent)) %>% 
  mutate(SynchComponent_r=recode_factor(SynchComponent, phi_S2C_L='Local_Synchony', phi_S2C_R='Regional_Synchrony')) %>% 
 ggplot()+aes(wNAO.window, value )+geom_point(size=2)+geom_smooth(method='lm')+
  facet_wrap(~SynchComponent_r)+theme(legend.position = 'right')+ylab('Community synchrony')+theme(axis.title = element_text(size=12))+xlab('Winter NAO window')
```




#Plot Fig.6
```{r fig.width=5, fig.height=5}
#library(patchwork)
plot_synchronies.skip /
  Plot_CVs.skip +plot_annotation(tag_levels = 'A')
```

#Plot of *Fig.6 Wangs decomposition* (ex fig.5)
```{r}
pdf('Fig.6_wangs_comp.pdf', w=5.5, h=4.5)
plot_synchronies.skip /
  Plot_CVs.skip +plot_annotation(tag_levels = 'A')

dev.off()
```



# Autocor Modles for community synchrony (local and regional)

```{r}
mod.phi_S2C_L=
  gls(phi_S2C_L~wNAO.window, data=wang_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),], cor=corAR1(0.6))

summary(mod.phi_S2C_L)

mod.phi_S2C_R=
  gls(phi_S2C_R~wNAO.window, data=wang_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),], cor=corAR1(0.8))
summary(mod.phi_S2C_R)
```


############################################################################
### Traits analyses for FDis calculation. Different options are presented.



#import the trait data
*this df is not normlised [0-1], will do this below*
```{r}
library(readxl)
Spp_subs_trt <- read_excel("Spp_subs_trt.xlsx", 
    sheet = "complete_upd")

names(Spp_subs_trt)[1]='taxa_id'

Spp_subs_trt=
Spp_subs_trt %>% 
  column_to_rownames(var='taxa_id')

```



# *order the trait df with same order of LB10.imp (same species order)
```{r}
Spp_subs_trt=
Spp_subs_trt[order(match(rownames(Spp_subs_trt), speciesID.imp)),]
  
```



# block for trait modelities
```{r}
block=c(10, 3, 5, 6, 6, 4, 6, 4, 3, 2, 2,2 )
```

# names of traits repeated
*this is useful for functions along traits*
```{r}
trt_names=rep(c('zonation','t_range','t_pref',  'feed', 'loc','drought', 'resis','size','diss','disp.cap', 'life.dur', 'lc' ), 
            block)

names(trt_names)=colnames(Spp_subs_trt)
```



######## ################################
## Version of functional diversity based on subset of (most complete) trait info
########################################

# check the completeness of trait info
```{r}
str(trt_names)
attributes(trt_names)

trt_names[c(31:40)] ## the drought & rsistence (6,7)
trt_names[c(45:49)] # dissemination & disp capacity (9,10)

trt_names[-c(31:40, 45:49)]

```

# update block
```{r}
block

block2=block[-c(6,7,9,10)]
```
# update the original trait info (subset)
```{r}
Spp_subs_trt

Spp_subs_trt2=
  Spp_subs_trt[,-c(31:40, 45:49)]

```

# update the fuzzy coded trait
```{r}
subs_trt_fuzzy2=
prep.fuzzy.var(Spp_subs_trt2, block2)
```

# export the subset of trait, fuzzy coded now
```{r}
write.csv(subs_trt_fuzzy2, 'subs_trt_fuzzy2.csv')
```


# Update rao.diversity and redundancy
```{r}
library(SYNCSA)
```

*the simpson rao is the null value based on random dissimilarities or where each spp is different in traits*
```{r}
rao.div2=
rao.diversity(LB10.imp[,-c(1,2)], subs_trt_fuzzy2)
```



# get the update rao diversity
```{r}
library(reshape2)

rao.div2=
cbind.data.frame(colsplit(rownames(as.data.frame(rao.div2$FunRao)), '_', names=c('Code', 'year')), rao.div2$FunRao, rao.div2$FunRedundancy)
names(rao.div2)[3]='FunRao'
names(rao.div2)[4]='FunRedundancy'

```


# the update moving window of functional diversity
```{r}
FunRao2.mw=as.data.frame(matrix(nrow=length(time.window5), ncol=10))
rownames(FunRao2.mw)=time.window5
colnames(FunRao2.mw)=lb10.siteID

for(s in 1: length(lb10.siteID)){
  
FunRao2.mw[,s]=rollmean(rao.div2 %>%  filter(Code==lb10.siteID[s]) %>% pull(FunRao), k=6, align = 'center')
}


# redundancy not used
FunRedun2.mw=as.data.frame(matrix(nrow=length(time.window5), ncol=10))
rownames(FunRedun2.mw)=time.window5
colnames(FunRedun2.mw)=lb10.siteID

for(s in 1: length(lb10.siteID)){
  
FunRedun.mw[,s]=rollmean(rao.div2 %>%  filter(Code==lb10.siteID[s]) %>% pull(FunRedundancy), k=6, align = 'center')
}


```

# the update plots of functinoal diversity
```{r}

plot_sync_rao=
cbind.data.frame(
synch.loreau.mw %>% pivot_longer(cols=c(2:11), values_to = 'CommSynch.Loreau'),
FunRao2.mw %>%  pivot_longer(cols=c(1:10), values_to = 'FunRao2', names_to = 'Code')) %>% 
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
  filter(time.window %in% time.window5_skip ) %>% 
   ggplot()+aes(FunRao2, sqrt(CommSynch.Loreau))+geom_point(aes(col=name))+geom_smooth(method='lm')+ylab('Community synchrony')+
  theme(axis.title = element_text(size=12))+scale_color_discrete(name='Stream')+xlab('Functional diversity')+geom_smooth(aes(col=name), se=F, method='lm', size=0.2)

plot_stab_rao=
cbind.data.frame(
comm.stab.mw %>% pivot_longer(cols=c(2:11), values_to = 'CommStab'),
FunRao2.mw %>%  pivot_longer(cols=c(1:10), values_to = 'FunRao2', names_to = 'Code')) %>% 
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
  filter(time.window %in% time.window5_skip ) %>% 
   ggplot()+aes(FunRao2, 1/CommStab)+geom_point(aes(col=name))+geom_smooth(method='lm', linetype='dashed')+ylab('Community variability')+
   theme(axis.title = element_text(size=12))+scale_color_discrete(name='Stream')+xlab('Functional diversity')+
  geom_smooth(aes(col=name), se=F, method='lm', size=0.2)


```


#
```{r fig.width=7, fig.height=4.5}
library(patchwork)

plot_sync_rao+plot_stab_rao + plot_layout(guides='collect') & theme(legend.position = 'top')

```


# *Plot Fig.7* as in the paper
```{r}
pdf('Fig.7_FD_synchrony.pdf', w=6.4, h=4.5)
plot_sync_rao+plot_stab_rao + plot_layout(guides='collect') & theme(legend.position = 'top')

dev.off()
```


# model for functional diversity vs synch & stability
Models that account for stream as random effect temporal autocorr
```{r}

m0.rao=
  cbind.data.frame(
synch.loreau.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>% 
  pivot_longer(cols=c(2:11), values_to = 'CommSynch.Loreau'),
FunRao2.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>%  
  pivot_longer(cols=c(1:10), values_to = 'FunRao', names_to = 'Code')) %>% 
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
  lm(CommSynch.Loreau~FunRao,  data=.)

# random stream effect
m.synch.rao=
cbind.data.frame(
synch.loreau.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>% 
  pivot_longer(cols=c(2:11), values_to = 'CommSynch.Loreau'),
FunRao2.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>%  
  pivot_longer(cols=c(1:10), values_to = 'FunRao', names_to = 'Code')) %>% 
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
   
  lme(CommSynch.Loreau~FunRao, random= ~1|Code, cor=corAR1(form=~as.numeric(as.factor(time.window))|Code), data=.)


summary(m.synch.rao)

# random slope too
m.synch.rao_b=
cbind.data.frame(
synch.loreau.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>% 
  pivot_longer(cols=c(2:11), values_to = 'CommSynch.Loreau'),
FunRao2.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>%  
  pivot_longer(cols=c(1:10), values_to = 'FunRao', names_to = 'Code')) %>% 
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
   
  lme(CommSynch.Loreau~FunRao, random= ~1+FunRao|Code, cor=corAR1(form=~as.numeric(as.factor(time.window))|Code), data=.)


AIC(m0.rao, m.synch.rao, m.synch.rao_b)


```



# Model for stability vs RoaQ, not significant
```{r}
m.stab.rao=
cbind.data.frame(
comm.stab.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>% 
  pivot_longer(cols=c(2:11), values_to = 'CommStab'),
FunRao2.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>%  
  pivot_longer(cols=c(1:10), values_to = 'FunRao', names_to = 'Code')) %>% 
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
   
  lme(CommStab~FunRao, random= ~1|Code, cor=corAR1(form=~as.numeric(as.factor(time.window))|Code), data=.)

summary(m.stab.rao)

```



## *Working on the functional deviation (null models for FDis)*


# function to shuffle names and provide a null FunRao value
```{r}

shuffle.names.rao=function(abb, trt){
  rownames(trt)=sample(rownames(trt), replace = F)
  SYNCSA::rao.diversity(abb, trt)$FunRao
}

# test funciton
shuffle.names.rao(LB10.imp[,-c(1,2)], subs_trt_fuzzy)

zio=subs_trt_fuzzy
rownames(zio)=sample(rownames(zio), replace = F)

rao.diversity(LB10.imp[,-c(1,2)],
zio)$FunRao
```

# combine the observed FunRao with replicates of null rao
```{r}
obs.null.rao2=
  cbind(rao.diversity(LB10.imp[,-c(1,2)], subs_trt_fuzzy2)$FunRao,
        replicate(300, shuffle.names.rao(LB10.imp[,-c(1,2)], subs_trt_fuzzy2)))


obs.null.rao=
  cbind(rao.diversity(LB10.imp[,-c(1,2)], subs_trt_fuzzy)$FunRao,
        replicate(300, shuffle.names.rao(LB10.imp[,-c(1,2)], subs_trt_fuzzy)))

```

# extract the deviations Rao and SES rao
```{r}

Rao_deviations=cbind.data.frame(code_year=rownames(obs.null.rao),
                                obs.Rao=obs.null.rao[,1],
                                Rao_dev=(obs.null.rao[,1]-rowMeans(obs.null.rao[,-1])),
                                SES.Rao=(obs.null.rao[,1]-rowMeans(obs.null.rao[,-1]) / apply(obs.null.rao[,-1],1, sd))



Rao_deviations2=cbind.data.frame(code_year=rownames(obs.null.rao2),
                                obs.Rao=obs.null.rao2[,1],
                                Rao_dev=(obs.null.rao2[,1]-rowMeans(obs.null.rao2[,-1])),
                                SES.Rao=(obs.null.rao2[,1]-rowMeans(obs.null.rao2[,-1]) / apply(obs.null.rao2[,-1],1, sd))
)

# add the Code info
Rao_deviations2$Code=rao.div$Code


```

```{r}
library(zoo)
library(tidyverse)
```

# calculate Rao deviation mowing window
```{r}
Rao_dev.mw=as.data.frame(matrix(nrow=length(time.window5), ncol=10))
rownames(Rao_dev.mw)=time.window5
colnames(Rao_dev.mw)=lb10.siteID

for(s in 1: length(lb10.siteID)){
  
Rao_dev.mw[,s]=rollmean(Rao_deviations %>%  filter(Code==lb10.siteID[s]) %>% pull(Rao_dev), k=6, align = 'center')
}
```

# Plot Fig.S4 - functional deviation
```{r fig.width=6}
cbind.data.frame(
synch.loreau.mw %>% pivot_longer(cols=c(2:11), values_to = 'CommSynch.Loreau'),
Rao_dev.mw %>%  pivot_longer(cols=c(1:10), values_to = 'Rao_deviation', names_to = 'Code')) %>% 
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
  filter(time.window %in% time.window5_skip) %>% 
   ggplot()+aes(Rao_deviation, CommSynch.Loreau)+geom_point(aes(col=name))+geom_smooth(method='lm', linetype='dashed')+ylab('Community synchrony')+
  theme(axis.title = element_text(size=12))+scale_color_discrete(name='Stream')+xlab('Functional deviation')+geom_smooth(aes(col=name), se=F, method='lm', size=0.2)+scale_color_brewer(name='Stream',palette='Paired')
```

#*plot Fig S4*
```{r}
pdf('Fig.S4_func_deviation.pdf', w=5,h=4)
cbind.data.frame(
synch.loreau.mw %>% pivot_longer(cols=c(2:11), values_to = 'CommSynch.Loreau'),
Rao_dev.mw %>%  pivot_longer(cols=c(1:10), values_to = 'Rao_deviation', names_to = 'Code')) %>% 
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
  filter(time.window %in% time.window5_skip) %>% 
   ggplot()+aes(Rao_deviation, CommSynch.Loreau)+geom_point(aes(col=name))+geom_smooth(method='lm', linetype='dashed')+ylab('Community synchrony')+
  theme(axis.title = element_text(size=12))+scale_color_discrete(name='Stream')+xlab('Functional deviation')+geom_smooth(aes(col=name), se=F, method='lm', size=0.2)+scale_color_brewer(name='Stream',palette='Paired')

dev.off()
```



```{r}
library(nlme)
```

*to update for functional deviation*
```{r}
m.dev=
cbind.data.frame(
synch.loreau.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>% 
  pivot_longer(cols=c(2:11), values_to = 'CommSynch'),
Rao_dev.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>%  
  pivot_longer(cols=c(1:10), values_to = 'Rao_deviation', names_to = 'Code')) %>% 
  
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
  lme(CommSynch~Rao_deviation, random= ~1|Code, cor=corAR1(0.6), data=.)


summary(m.dev)

m.dev2=
cbind.data.frame(
synch.loreau.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>% 
  pivot_longer(cols=c(2:11), values_to = 'CommSynch'),
Rao_dev2.mw [c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>%  
  pivot_longer(cols=c(1:10), values_to = 'Rao_deviation', names_to = 'Code')) %>% 
  
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
  lme(CommSynch~Rao_deviation, random= ~1|Code, cor=corAR1(0.6), data=.)

summary(m.dev2)
```


