---
title: "Stability_partit"
author: "Stefano Larsen"
date: "10/14/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 6.5)
```

```{r}
theme_set(theme_bw())
```


# This script holds some codes for running partition of stability as Wang & Loreau 2019 (Lise's take)


# This is the original Wang's code for partitioning
```{r}
var.partition <-function(metacomm_tsdata){
## The function "var.partition" performs the partitioning of variability 
## across hierarchical levesl within a metacommunity.
## The input array "metacomm_tsdata" is an N*T*M array. The first dimension represents N species, 
## the second represents time-series observations of length T, and the third represents M local communities. 
## The output includes four variability and four synchrony metrics as defined in the main text.
## Note that, to be able to handle large metacommunities, this code has avoided calculating all covariance.

ts_metacom <-apply(metacomm_tsdata,2,sum,na.rm=T)
ts_patch <-apply(metacomm_tsdata,c(2,3),sum,na.rm=T)
ts_species <-apply(metacomm_tsdata,c(1,2),sum,na.rm=T)
sd_metacom <-sd(ts_metacom,na.rm=T)
sd_patch_k <-apply(ts_patch,2,sd,na.rm=T)
sd_species_i <-apply(ts_species,1,sd,na.rm=T)
sd_species_patch_ik <-apply(metacomm_tsdata,c(1,3),sd,na.rm=T)
mean_metacom <-mean(ts_metacom,na.rm=T)

CV_S_L <-sum(sd_species_patch_ik,na.rm=T)/mean_metacom  #local-scale average species variability (weighted average of local                                                             population variability across sites and species)
CV_C_L <-sum(sd_patch_k,na.rm=T)/mean_metacom            #local-scale average community variability (weighted average of community                                                             variability across sites)
CV_S_R <-sum(sd_species_i,na.rm=T)/mean_metacom          #regional-scale average species variability (weighted average of                                                                       metapopulation variability across species)
CV_C_R <-sd_metacom/mean_metacom                  #metacommunity variability (CV total metacommunity biomass)

phi_S_L2R <-CV_S_R/CV_S_L  #species-level spatial synchrony (weighted average intraspecific across species)
phi_C_L2R <-CV_C_R/CV_C_L  #community-level spatial synchrony  (of total community biomass across sites)
phi_S2C_L <-CV_C_L/CV_S_L  #local-scale species synchrony     (weighted average interspecific across patches)
phi_S2C_R <-CV_C_R/CV_S_R   #regional-scale species synchrony  (among different meta-populations [sum of covariances                                                                                                           beween species]
partition_3level <-c(CV_S_L=CV_S_L, CV_C_L=CV_C_L, CV_S_R=CV_S_R, CV_C_R=CV_C_R, phi_S_L2R=phi_S_L2R, phi_C_L2R=phi_C_L2R, phi_S2C_L=phi_S2C_L, phi_S2C_R=phi_S2C_R)
return(partition_3level)
}
```


## Convert the bio data into an array to feed the var.partition function. Will use the imputed data this time on LB10



# gather bugs data in long format -useful for subsequent manipulations
```{r}
bugs.long.imp=
LB10.imp %>% 
  gather(Species, abund, -c(Code, year))
```

# code, year and species to factor
```{r}
bugs.long.imp$Code=as.factor(bugs.long.imp$Code)
bugs.long.imp$year=as.factor(bugs.long.imp$year)
bugs.long.imp$Species=as.factor(bugs.long.imp$Species)
```

# define the number of years, sites and species overall
```{r}
nlevels(bugs.long.imp$year)
nlevels(bugs.long.imp$Code)
nlevels(bugs.long.imp$Species)
```

# create vector keeping year, site and species ID - useful for working in loops
```{r}
yearID=sort(unique(bugs.long$year))
codeID=sort(unique(bugs.long$Code))
speciesID=sort(unique(bugs.long$Species))
```

# Create a list keeping matrix of species x year - each spp a timeseries for each stream
```{r}
mylist=list()
```

```{r}

for(code in 1:nlevels(bugs.long.imp$Code)){
  yearID=sort(unique(bugs.long.imp$year))
  codeID=sort(unique(bugs.long.imp$Code))
  speciesID=sort(unique(bugs.long.imp$Species))
  bugs.long.code=bugs.long.imp[bugs.long.imp$Code==codeID[code],]
  bugs.long.mat.code=(spread(bugs.long.code[,-1], year, abund, drop=F, fill=0))
   rownames(bugs.long.mat.code)=bugs.long.mat.code[,1]
   bugs.long.mat.code[,1]=NULL
 mylist[[code]]=bugs.long.mat.code
}
```

#  see the list
```{r}
mylist[[2]]
```

# Convert the list of species x year (for each stream) into an *ARRAY* with dim = species x years x site
*this could be useful for the CV stability function of Wang*

```{r}
myarray=array(as.numeric(unlist(mylist)), dim=c(nlevels(bugs.long.imp$Species), nlevels(bugs.long.imp$year), nlevels(bugs.long.imp$Code)))
dimnames(myarray)=list(speciesID, yearID, codeID)

myarray[,,1]
myarray[,,3]

dim(myarray)# 65 species, 34 years, 10 sites (with no holes, i.e. imputed)
```





# *Wang's partition of stability*
# also cqn be used with moving windo
```{r}
res.wang.part=var.partition(myarray)


dim(myarray)

var.partition(myarray)
```

```{r}


(res.wang.part[1])*(res.wang.part[7])*(res.wang.part[6])


log10(res.wang.part[1]) + log10(res.wang.part[7])+log10(res.wang.part[6])

# equals
log10(res.wang.part[4])
```

log(0.7937204)



#######################
# Derive for each stream the stability metrics namely community stability and synchrony (codyn)
```{r}
stability.metrics=as.data.frame(matrix(ncol=3, nrow=10))
rownames(stability.metrics)=lb10.siteID
colnames(stability.metrics)=c('comm.stab', 'synch.loreau', 'sycnh.gross')

```

# Fill the stability metric for each site
```{r}

# communty stability
for (s in 1:length(lb10.siteID)){
stability.metrics[s,1]=
 LB10.imp %>% 
  filter(Code==lb10.siteID[s]) %>% 
  pivot_longer(col=c(3:67)) %>% 
  select(-Code) %>% 
  mutate(name=as.factor(name), year=as.integer(year)) %>% 
 community_stability (time.var = 'year',  abundance.var = 'value')
}

# comm synchonries

for (s in 1:length(lb10.siteID)){
stability.metrics[s,2]=
 LB10.imp %>% 
  filter(Code==lb10.siteID[s]) %>% 
  pivot_longer(col=c(3:67)) %>% 
  select(-Code) %>% 
  mutate(name=as.factor(name), year=as.integer(year)) %>% 
 synchrony (time.var = 'year', species.var='name', abundance.var = 'value',metric = 'Loreau' )
}


for (s in 1:length(lb10.siteID)){
stability.metrics[s,3]=
 LB10.imp %>% 
  filter(Code==lb10.siteID[s]) %>% 
  pivot_longer(col=c(3:67)) %>% 
  select(-Code) %>% 
  mutate(name=as.factor(name), year=as.integer(year)) %>% 
 synchrony (time.var = 'year', species.var='name', abundance.var = 'value',metric = 'Gross' )
}



```

# Plot stability metrics and diversity
```{r}
cbind.data.frame(lb10.div.means[,-1], stability.metrics) %>% 
pairs(panel=(panel.smooth))
```


## *Nest the lb10.imp over each site*
```{r}
LB10.imp.n=LB10.imp %>% 
  group_by(Code) %>% 
  nest()
```




# *Stability function* to run over moving window of n years
```{r}

# this function only works within nested version of lb10 nested for sites

stabf=function(df){
df %>% 
  #filter(Code==lb10.siteID[1]) %>% 
  pivot_longer(col=speciesID.imp) %>% # remove also 
  #select(-Code) %>% 
  mutate(name=as.factor(name), year=as.integer(year)) %>% 
 community_stability (time.var = 'year',  abundance.var = 'value')
}


stabf(b.CI1.imp)
```

#tests
```{r}

b.CI1.imp %>% 
  pivot_longer(col=speciesID.imp)

LB10.imp.n$data[[1]] %>% 
  pivot_longer(col=speciesID.imp)

# the stability funciton b works on both the nested and original site data
stabf(b.CI1.imp) # 

stabf(LB10.imp.n$data[[1]])

```



# run the stability function on each nested site
```{r}
LB10.imp.n=
LB10.imp.n %>% 
  mutate(com.stability=unlist(map(data, stabf)))


LB10.imp.n$data[[1]]
```



# function that claculate the comm stability metric *for a window of 5 years (moving by 1y each time)*
```{r}

# test for creating the function
zio=list()
ywindow=5
for(i in 1:(nrow(b.CI1.imp)-ywindow)){
zio[[i]]=  stabf.b(b.CI1.imp[i:(i+4),])
}



stabf.mw=function(df){
zio=list()
ywindow=5
for(i in 1:(nrow(df)-ywindow)){
zio[[i]]=  stabf(df[i:(i+5),])}
return(unlist(zio))
}

stabf.mw(b.CI1.imp)

stabf.mw(LB10.imp.n$data[[1]])

```

```{r}
stabf.mw(b.LI7.imp)
```




# run the moving window stability function on each site (nested)
```{r}

LB10.imp.n =
LB10.imp.n %>% 
  mutate(stability.mw=(map(data, stabf.mw)))

LB10.imp.n$stability.mw[[1]]

rbindlist(list(LB10.imp.n$stability.mw))

```
#extract the moving window of comm stability into a df
```{r}

time.window5=matrix(nrow=29)
for(i in 1:29){
time.window5[i,]=paste(time.lb10[i], time.lb10[i+5], sep='_')  
}

comm.stab.mw=cbind.data.frame(time.window=time.window5,
  rbindlist(list(LB10.imp.n$stability.mw))
)

colnames(comm.stab.mw)[2:11]=rownames(stability.metrics)
```

# Plot the comm stability
```{r fig.width=9}
comm.stab.mw %>% 
  pivot_longer(cols = -1) %>% 
  ggplot()+aes(time.window, value, group=1)+geom_path()+facet_wrap(~name)+theme(axis.text.x = element_text(angle=90))
```




# *Synchrony function* to run over moving window of n years
```{r}

# this function only works within nested version of lb10 nested for sites

synchf=function(df){
df %>% 
  #filter(Code==lb10.siteID[1]) %>% 
  pivot_longer(col=speciesID.imp) %>% # remove also 
  #select(-Code) %>% 
  mutate(name=as.factor(name), year=as.integer(year)) %>% 
 synchrony (time.var = 'year',species.var = 'name',  abundance.var = 'value', metric = 'Loreau')
}


synchf(b.CI1.imp)
```

# synchrony funciton over 5 y moving window (as done fo stability)
```{r}

synchf.mw=function(df){
zio=list()
ywindow=5
for(i in 1:(nrow(df)-ywindow)){
zio[[i]]=  synchf(df[i:(i+5),])}
return(unlist(zio))
}
```



# run the moving window stability function on each site (nested)
```{r}

LB10.imp.n =
LB10.imp.n %>% 
  mutate(synch.loreau.mw=(map(data, synchf.mw)))

LB10.imp.n$synch.loreau.mw[[1]]

rbindlist(list(LB10.imp.n$stability.mw))
```

# extractt he synchrony moving window loreau
```{r}
synch.loreau.mw=cbind.data.frame(time.window=time.window5,
  rbindlist(list(LB10.imp.n$synch.loreau.mw))
)

colnames(synch.loreau.mw)[2:11]=rownames(stability.metrics)
```


# Plot the comm synchrony
```{r fig.width=9}
synch.loreau.mw %>% 
  pivot_longer(cols = -1) %>% 
  ggplot()+aes(time.window, value, group=1)+geom_path()+facet_wrap(~name)+theme(axis.text.x = element_text(angle=90))+ylab('Community synchrony')
```


```{r}
plotmag(
wpmf(
cleandat(as.matrix(t(synch.loreau.mw[,-1])), times=c(1:29), clev = 4)$cdat,
times=c(1:29), sigmethod = 'quick'))
```




# Get the 6y timewindow for nao
```{r}
wNAO_window=
cbind.data.frame(
  time.window5,
NAOwindow=rollmean(NAOseason %>% filter(Season=='Winter') %>% filter(winter.year %in% c(1985:2018)) %>%  pull(mean.nao), k=6, align = 'center'))

```



```{r}
cbind.data.frame(NAOwindow=wNAO_window$NAOwindow, synch.loreau.mw) %>% 
pivot_longer(cols=-c(1,2))  %>% 
  ggplot()+aes(NAOwindow, value)+geom_point()+geom_smooth(method='lm')+facet_wrap(~name)

```

```{r}
cbind.data.frame(NAOwindow=wNAO_window$NAOwindow, comm.stab.mw) %>% 
pivot_longer(cols=-c(1,2))  %>% 
  ggplot()+aes(NAOwindow, value)+geom_point()+geom_smooth(method='lm')+facet_wrap(~name, scales = 'free')
```
# Plot the maen synchrony and stability vs nao
```{r}
cbind.data.frame(wNAO_window, mean.stability=rowMeans(comm.stab.mw[,-1])) %>% 
  ggplot()+aes(NAOwindow, mean.stability, label=time.window5)+geom_point()+geom_text_repel(size=3)+geom_smooth(method='lm')



cbind.data.frame(wNAO_window, mean.synchrony=rowMeans(synch.loreau.mw[,-1])) %>% 
  ggplot()+aes(NAOwindow, mean.synchrony, label=time.window5)+geom_point()+geom_text_repel(size=3)+geom_smooth(method='lm')

```
# Mean stability vs NAO for each stream, with color by mean richness
```{r}
cbind.data.frame(wNAO_window, comm.stab.mw) %>% 
  pivot_longer(cols=c(4:13), names_to = 'Code') %>% left_join(lb10.div.means, by='Code' ) %>% 
  ggplot()+aes(NAOwindow, value, label=time.window5)+geom_point(aes(col=mrich))+geom_smooth(method='lm')+
  scale_color_gradient(high='darkred', low='gold3')+facet_wrap(~Code, scale='free')

```
# Functions to extract the effect of NAO on stability and synchrony (as slope on mw data)
```{r}
CommStabFn=function(df){
  coefficients(lm(CommStab~NAOwindow, df))[2]
}

CommSynchFn=function(df){
  coefficients(lm(CommSynchLoreau~NAOwindow, df))[2]
}
```

# Plot the effect of NAO on stability (as slope) vs the mean richness of streams. No clear pattern
```{r}
cbind.data.frame(wNAO_window, comm.stab.mw) %>% 
  pivot_longer(cols=c(4:13), names_to = 'Code', values_to = 'CommStab') %>% left_join(lb10.div.means, by='Code' ) %>% 
  group_by(Code) %>% nest() %>% 
  mutate(StabEffect=map(data, CommStabFn)) %>% unnest() %>% 
  ggplot()+aes(mrich,StabEffect)+geom_point()+geom_smooth()
```

```{r}
cbind.data.frame(wNAO_window, comm.stab.mw) %>% 
  pivot_longer(cols=c(4:13), names_to = 'Code', values_to = 'CommStab') %>% left_join(rao.div.means, by='Code' ) %>% 
  group_by(Code) %>% nest() %>% 
  mutate(StabEffect=map(data, CommStabFn)) %>% unnest() %>% 
  ggplot()+aes(mean.FunRao,StabEffect)+geom_point()+geom_smooth()

cbind.data.frame(wNAO_window, synch.loreau.mw) %>% 
  pivot_longer(cols=c(4:13), names_to = 'Code', values_to = 'CommSynchLoreau') %>% left_join(rao.div.means, by='Code' ) %>% 
  group_by(Code) %>% nest() %>% 
  mutate(SynchEffect=map(data, CommSynchFn)) %>% unnest() %>% 
  ggplot()+aes(mean.FunRao,SynchEffect)+geom_point()+geom_smooth()

```




# Eventually calculate the richness (mean site or metacom) for each 6-y time window



```{r}
Richness.mw=as.data.frame(matrix(nrow=length(time.window5), ncol=10))
rownames(Richness.mw)=time.window5
colnames(Richness.mw)=lb10.siteID

for(s in 1: length(lb10.siteID)){
  
Richness.mw[,s]=rollmean(lb10.div %>%  filter(Code==lb10.siteID[s]) %>% pull(richness), k=6, align = 'center')
}

```

# *not particular effect of mean sstream richness on the 6y moving window of stability or synchrony*

```{r fig.width=6}


cbind.data.frame(wNAO_window, mean.rich.mw=rowMeans(Richness.mw),
                 mean.synch=rowMeans(synch.loreau.mw [,-1])) %>% 
  ggplot()+aes(NAOwindow, mean.synch, label=time.window5)+geom_point()+geom_text_repel(aes(col=mean.rich.mw), size=2.5)+
  geom_point(aes(col=mean.rich.mw), size=3)+geom_smooth(method='lm')+
  scale_color_gradient(high='darkred', low='gold3')+ylab('Mean community synchrony (Loreau)')


cbind.data.frame(wNAO_window, mean.rich.mw=rowMeans(Richness.mw),
                 mean.stability=rowMeans(comm.stab.mw[,-1])) %>% 
  ggplot()+aes(NAOwindow, mean.stability, label=time.window5)+geom_text_repel(aes(col=mean.rich.mw), size=2.5)+
  geom_point(aes(col=mean.rich.mw), size=3)+
  geom_smooth(method='lm', col='grey50', alpha=.2)+
  scale_color_gradient(high='darkred', low='gold3')+ylab('Mean community stability')



```

# No strong relation between Loreau synchrony and richness
```{r}
cbind.data.frame(wNAO_window, mean.rich.mw=rowMeans(Richness.mw),
                 mean.stability=rowMeans(comm.stab.mw[,-1])) %>% 
  ggplot()+aes(mean.rich.mw, mean.stability)+geom_point()

cbind.data.frame(wNAO_window, mean.rich.mw=rowMeans(Richness.mw),
                 mean.synch=rowMeans(synch.loreau.mw[,-1])) %>% 
  ggplot()+aes(mean.rich.mw, mean.synch)+geom_point()

```

#Regional scale richness and shannon and its moving window over 6y
```{r}
reg.richness=
LB10.imp[,-1] %>% group_by(year) %>% summarise_all(sum) %>% select(-1) %>% apply(1,renyi, scales=0, hill=T)

reg.richness %>% rollmean(k=6, align = 'center')



reg.shannon=
LB10.imp[,-1] %>% group_by(year) %>% summarise_all(sum) %>% select(-1) %>% apply(1,renyi, scales=1, hill=T)

```

# Again no strong effect of regional richness diversity on mean community stability as affected by wNAO moving window
```{r}
cbind.data.frame(wNAO_window, reg.rich.mw=reg.richness %>% rollmean(k=6, align = 'center'),
                 mean.synch=rowMeans(synch.loreau.mw [,-1])) %>% 
  ggplot()+aes(NAOwindow, mean.synch, label=time.window5)+geom_point()+geom_point(aes(col=reg.rich.mw), size=3)+geom_smooth(method='lm')+
  scale_color_gradient(high='darkred', low='gold3')

cbind.data.frame(wNAO_window, reg.rich.mw=reg.richness %>% rollmean(k=6, align = 'center'),
                 mean.stability=rowMeans(comm.stab.mw[,-1])) %>% 
  ggplot()+aes(NAOwindow, mean.stability, label=time.window5)+geom_point()+geom_point(aes(col=reg.rich.mw), size=3)+geom_smooth(method='lm')+
  scale_color_gradient(high='darkred', low='gold3')

cbind.data.frame(wNAO_window, reg.shan.mw=reg.shannon %>% rollmean(k=6, align = 'center'),
                 mean.stability=rowMeans(comm.stab.mw[,-1])) %>% 
  ggplot()+aes(NAOwindow, mean.stability, label=time.window5)+geom_point()+geom_point(aes(col=reg.shan.mw), size=3)+geom_smooth(method='lm')+
  scale_color_gradient(high='darkred', low='gold3')

```
### Quick test of wang decmposition using moving window

```{r}
wang_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=length(res.wang.part)))
rownames(wang_mw)=time.window5
colnames(wang_mw)=names(res.wang.part)

var.partition(myarray[,c(1: (1+5)),])


for(y in 1: length(time.window5)){
 wang_mw[y,]= var.partition(myarray[,c(y: (y+5)),])
}

```
# add wNAOwindow on the wang's stability window
```{r}
wang_mw$wNAO.window=wNAO_window$NAOwindow
```

# *wNAO affects all components of Wang stability*
```{r}
wang_mw %>% 
  pivot_longer(cols=c(1:8)) %>% 
  ggplot()+aes(wNAO.window, value)+geom_point()+geom_smooth(method='lm')+facet_wrap(~name, scales = 'free')
```
# breaking down the CV and synch effect across hierarchies
```{r}
res.wang.part

# CV_C_R= CV_S_R * phi_S2C_R   ; this is the 1st decompsitiono of metacom CV

res.wang.part[3]*res.wang.part[8]

```

# similar effect of NAO on CV_metapopulation and regional scale spp synchorny
```{r}
mod_CV_S_R=
(lm(log10(wang_mw$CV_S_R)~wang_mw$wNAO.window))

mod_phi_S2C_R=
  lm(log10(wang_mw$phi_S2C_R)~wang_mw$wNAO.window)

# stadard coeff CV_S_R
mod_CV_S_R$coefficients[2]*(sd(wang_mw$wNAO.window)/
                              sd(wang_mw$CV_S_R))
# stand coeff Phi_S2C_R
mod_phi_S2C_R$coefficients[2]* (sd(wang_mw$wNAO.window)/
                                  sd(wang_mw$phi_S2C_R))

sd(wang_mw$CV_S_R)#sd.y
sd(wang_mw$phi_S2C_R)
sd(wang_mw$wNAO.window)#sd.x




```

```{r}
wang_mw[c(1,8,9)] %>% 
  pivot_longer(cols=c(1:2)) %>% 
  #mutate(name=factor(name, levels=c('CV_S_L', 'CV_S_R', 'CV_C_R'))) %>% 
  ggplot()+aes(wNAO.window, value, col=name)+geom_point()+geom_line()+  scale_y_log10()+geom_smooth(method='lm')
```




```{r fig.width=5.5, fig.height=4.4}
log10(res.wang.part[1]) + log10(res.wang.part[7])+log10(res.wang.part[6])

# equals
log10(res.wang.part[4])

wang_mw[c(1,3,4,9)] %>% 
  pivot_longer(cols=c(1:3)) %>% 
  mutate(name=factor(name, levels=c('CV_S_L', 'CV_S_R', 'CV_C_R'))) %>% 
  ggplot()+aes(wNAO.window, value, col=name)+geom_point(size=2)+geom_line(size=1.2)+  scale_y_log10()+
  
  geom_smooth(method='lm', se=F)+scale_color_discrete(name='Components', values=c("darkred", 'forestgreen', 'grey20'), labels=c('CV population', 'CV metapopulation', 'CV metacommunity'))+theme(legend.position = 'top')+ylab('Variability (log10)')+theme(axis.title = element_text(size=12))
  
  scale_color_manual(values=c("darkred", 'forestgreen', 'grey20'))+theme(legend.title = element_blank())

```
```{r fig.width=5.5, fig.height=4.4}
wang_mw[c(1,3,4,9)] %>% 
  pivot_longer(cols=c(1:3)) %>% 
  mutate(name=factor(name, levels=c('CV_S_L', 'CV_S_R', 'CV_C_R'))) %>% 
  ggplot()+aes(wNAO.window, value, col=name)+geom_point(size=2)+geom_line(size=1.2)+  scale_y_log10()+
  
  geom_smooth(method='lm', se=F)+theme(legend.position = 'top')+ylab('Variability (log10)')+theme(axis.title = element_text(size=12))+
  scale_color_manual(values=c("orange", 'forestgreen', 'grey20'), labels=c('CV population', 'CV metapopulation', 'CV metacommunity'))+theme(legend.title = element_blank())

  
 
```




# Wang components of variability and synchrony (non-spatial) vs wNAO.mw color code by functional Rao
```{r fig.width=5.5}
cbind.data.frame(wang_mw, FunRao.mw=rowMeans(FunRao.mw)) %>% 
  pivot_longer(cols=c(1,2,3,4), names_to = 'StabComponent') %>% 
  ggplot()+aes(wNAO.window, value )+geom_point(aes(col=FunRao.mw), size=3)+geom_smooth(method='lm')+
  scale_color_gradient(high='darkred', low='gold3')+facet_wrap(~StabComponent)+theme(legend.position = 'right')+ylab('Variability')

cbind.data.frame(wang_mw, FunRao.mw=rowMeans(FunRao.mw)) %>% 
  pivot_longer(cols=c(7,8), names_to = 'SynchComponent') %>% 
  ggplot()+aes(wNAO.window, value )+geom_point(aes(col=FunRao.mw), size=3)+geom_smooth(method='lm')+
  scale_color_gradient(high='darkred', low='gold3')+facet_wrap(~SynchComponent)+theme(legend.position = 'right')+ylab('(Meta)Community synchrony')


```
# Plot of community synchorny (phi) at local and regional scale
```{r}
cbind.data.frame(wang_mw, FunRao.mw=rowMeans(FunRao.mw)) %>% 
  pivot_longer(cols=c(7,8), names_to = 'SynchComponent') %>% 
  mutate(SynchComponent=factor(SynchComponent)) %>% 
  mutate(SynchComponent_r=recode_factor(SynchComponent, phi_S2C_L='Local_Synchony', phi_S2C_R='Regional_Synchrony')) %>% 
  ggplot()+aes(wNAO.window, value )+geom_point(size=3)+geom_smooth(method='lm')+
  facet_wrap(~SynchComponent_r)+theme(legend.position = 'right')+ylab('Community synchrony')+theme(axis.title = element_text(size=12))
```


# apprent effect of functional diversity on community synchrony is not significant when looking at overll mean
```{r fig.width=5}
cbind.data.frame(wang_mw, FunRao.mw=rowMeans(FunRao.mw)) %>% 
  pivot_longer(cols=c(7,8), names_to = 'SynchComponent') %>% 
  ggplot()+aes(FunRao.mw,value)+geom_point()+facet_wrap(~SynchComponent)+geom_smooth(method = 'lm')+ylab('Meta)Community synchrony')

cbind.data.frame(wang_mw, FunRao.mw=rowMeans(FunRao.mw)) %>% 
  select(FunRao.mw, phi_S2C_R) %>% lm(phi_S2C_R~FunRao.mw, data=.) %>% summary()

```

# plots Comm synchrony and stability vs FunRao, colored by site
```{r}
plot1=
cbind.data.frame(
synch.loreau.mw %>% pivot_longer(cols=c(2:11), values_to = 'CommSynch.Loreau'),
FunRao.mw %>%  pivot_longer(cols=c(1:10), values_to = 'FunRao', names_to = 'Code')) %>% 
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
   ggplot()+aes(FunRao, CommSynch.Loreau)+geom_point(aes(col=name))+geom_smooth(method='lm')+ylab('Community synchrony')+
  theme(axis.title = element_text(size=12))+scale_color_discrete(name='Stream')

plot2=
cbind.data.frame(
comm.stab.mw %>% pivot_longer(cols=c(2:11), values_to = 'CommStab'),
FunRao.mw %>%  pivot_longer(cols=c(1:10), values_to = 'FunRao', names_to = 'Code')) %>% 
  left_join(wNAO_window, by=c('time.window' = 'time.window5')) %>% 
   ggplot()+aes(FunRao, 1/CommStab)+geom_point(aes(col=name))+geom_smooth(method='lm')+ylab('Community variation')+
   theme(axis.title = element_text(size=12))+scale_color_discrete(name='Stream')


```


```{r fig.width=7, fig.height=4.5}
#library(patchwork)

plot1+plot2 + plot_layout(guides='collect') & theme(legend.position = 'top')

```



```{r fig.width=5.5}
cbind.data.frame(wang_mw, MeanRich.mw=rowMeans(Richness.mw)) %>% 
  pivot_longer(cols=c(1,2,3,4), names_to = 'StabComponent') %>% 
  ggplot()+aes(wNAO.window, value )+geom_point(aes(col=MeanRich.mw), size=3)+geom_smooth(method='lm')+
  scale_color_gradient(high='darkred', low='gold3')+facet_wrap(~StabComponent)+theme(legend.position = 'right')+ylab('Variability')

cbind.data.frame(wang_mw, MeanRich.mw=rowMeans(Richness.mw)) %>% 
  pivot_longer(cols=c(7,8), names_to = 'SynchComponent') %>% 
  ggplot()+aes(wNAO.window, value )+geom_point(aes(col=MeanRich.mw), size=3)+geom_smooth(method='lm')+
  scale_color_gradient(high='darkred', low='gold3')+facet_wrap(~SynchComponent)+theme(legend.position = 'right')+ylab('(Meta)Community synchrony')
```





# Calculate richness synchrony using moving window 6y

```{r}
rich.synch_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=1))


rownames(rich.synch_mw)=time.window5
colnames(rich.synch_mw)='Richness.synch.mw'



# the richness synch as simple cor
for(y in 1: length(time.window5)){
 rich.synch_mw[y,]= mean(
   cor(
     t(richness.ts[,c(y:(y+5))]), method='spearman'
     )
   )
}
```

# add the NAO window inot the richness synch moving window
```{r}
rich.synch_mw$wNAOwindow=wNAO_window$NAOwindow[match(rownames(rich.synch_mw), wNAO_window$time.window5)]


```

# NAO window does affect richness synchrony (higher synch with higher NAO), but relation is weak
*if we want to limit temporal autocor, just use every second record for the analysis. Limit overlap across mw*
```{r fig.width=5}
rich.synch_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>% #with every second year-window only
  ggplot()+aes(wNAOwindow, Richness.synch.mw, label=row.names(rich.synch_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),]))+geom_point(size=2)+geom_text_repel(size=2.5)+
  geom_smooth(method='lm', col='grey50', alpha=.3)+ylab('Richness synchrony mw')+ xlab('Winter NAO mw')


summary(lm(Richness.synch.mw~wNAOwindow, data=rich.synch_mw))
```

```{r}
m1=lm(Richness.synch.mw~wNAOwindow, data=rich.synch_mw)
summary(m1)

pacf(residuals(m1))

library(nlme)
m2=gls(Richness.synch.mw~wNAOwindow, data=rich.synch_mw, cor=corAR1(0.8))

summary(m2)



m2b=gls(Richness.synch.mw~wNAOwindow, data=rich.synch_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),], cor=corAR1(0.6))
summary(m2b)

c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),]
```


```{r}
library(zoo)
library(ggrepel)
```
```{r}

```


```{r fig.width=6}
rich.synch_mw %>% 
  mutate(RegRich.mw= reg.richness %>% rollmean(k=6, align = 'center')) %>% 
  ggplot()+aes(wNAOwindow, Richness.synch.mw, label=row.names(rich.synch_mw))+geom_point(size=2, aes(col=RegRich.mw))+geom_text_repel(size=2.5)+ scale_color_gradient(high='darkred', low='gold3')+
  geom_smooth(method='lm', col='grey50', alpha=.3)+ylab('Richness synchrony mw')+ xlab('Winter NAO mw')


rich.synch_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>% 
  mutate(MeanRich.mw= rowMeans(Richness.mw)) %>% 
  ggplot()+aes(wNAOwindow, Richness.synch.mw, label=row.names(rich.synch_mw))+geom_point(size=2, aes(col=MeanRich.mw))+geom_text_repel(size=3)+ scale_color_gradient(high='darkred', low='gold3')+
  geom_smooth(method='lm', col='grey50', alpha=.3)+ylab('Richness synchrony mw')+ xlab('Winter NAO mw')

```





# Claculate non-param function NCF as moving window
```{r}

rich.ncf_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=1))


rownames(rich.ncf_mw)=time.window5
colnames(rich.ncf_mw)='rich.cbar_interc.mw'



# the richness synch as simple cor
for(y in 1: length(time.window5)){
 rich.ncf_mw[y,]= 
   Sncf(x=coordsLB10$X, y=coordsLB10$Y, z=as.matrix((richness.ts[,c(y:(y+5))])), type = 'perm', resamp = 500)$real$cbar
   
}
```
```{r}
rich.ncf_mw$wNAOwindow=wNAO_window$NAOwindow[match(rownames(rich.ncf_mw), wNAO_window$time.window5)]
```

```{r}
rich.ncf_mw %>% 
  ggplot()+aes(wNAOwindow, rich.cbar_interc.mw, label=row.names(rich.synch_mw))+geom_point(size=2)+geom_text_repel(size=2.5)+
  geom_smooth(method='lm', col='grey50', alpha=.3)+ylab('Richness ncf mw')+ xlab('Winter NAO mw')


summary(lm(rich.ncf_mw$rich.cbar_interc.mw~rich.ncf_mw$wNAOwindow))

```



# Calculate Simpson synchrony moving window
```{r}
simp.synch_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=1))


rownames(simp.synch_mw)=time.window5
colnames(simp.synch_mw)='Simpson.synch.mw'



# the richness synch as simple cor
for(y in 1: length(time.window5)){
 simp.synch_mw[y,]= mean(
   cor(
     t(simpson.ts[,c(y:(y+5))]), method='spearman'
     )
   )
}
```


```{r}
simp.synch_mw$wNAOwindow=wNAO_window$NAOwindow[match(rownames(simp.synch_mw), wNAO_window$time.window5)]
```

# NAO window does NOT affect Simpson synchrony 
```{r}
simp.synch_mw %>% 
  ggplot()+aes(wNAOwindow, Simpson.synch.mw)+geom_point(size=3)+geom_smooth(method='lm')



```



# Calculate Abundance synchrony moving window
```{r}
abund.synch_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=1))


rownames(abund.synch_mw)=time.window5
colnames(abund.synch_mw)='Abund.synch.mw'



# the richness synch as simple cor
for(y in 1: length(time.window5)){
 abund.synch_mw[y,]= mean(
   cor(
     t(abund.ts[,c(y:(y+5))]), method='spearman'
     )
   )
}
```

```{r}
abund.synch_mw$wNAOwindow=wNAO_window$NAOwindow[match(rownames(abund.synch_mw), wNAO_window$time.window5)]
```

# NAO window does affect synchrony in abundance. 
```{r fig.width=6}
abund.synch_mw %>% 
  ggplot()+aes(wNAOwindow, Abund.synch.mw)+geom_point(size=3)+geom_smooth(method='lm')

abund.synch_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),] %>% 
  #mutate(RegRich.mw= reg.richness %>% rollmean(k=6, align = 'center')) %>% 
  #mutate(MeanRich.mw= rowMeans(Richness.mw)) %>% 
  ggplot()+aes(wNAOwindow, Abund.synch.mw, label=row.names(abund.synch_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),]))+geom_point(size=2)+geom_text_repel(size=3)+
  
  geom_smooth( col='grey50', alpha=.3, method='lm')+ylab('Abundance synchrony mw')+ xlab('Winter NAO mw')



abund.synch_mw %>% 
  #mutate(RegRich.mw= reg.richness %>% rollmean(k=6, align = 'center')) %>% 
  mutate(MeanRich.mw= rowMeans(Richness.mw)) %>% 
  ggplot()+aes(wNAOwindow, Abund.synch.mw, label=row.names(abund.synch_mw))+geom_point(size=2, aes(col=MeanRich.mw))+geom_text_repel(size=3)+
  scale_color_gradient(high='darkred', low='gold3')+
  geom_smooth( col='grey50', alpha=.3)+ylab('Abundance synchrony mw')+ xlab('Winter NAO mw')


```

```{r}
m3=lm(abund.synch_mw$Abund.synch.mw~ abund.synch_mw$wNAOwindow)
summary(m3)
acf(residuals(m3))

pacf(residuals(m3))

m4=gls(Abund.synch.mw~wNAOwindow,data=abund.synch_mw, cor=corAR1(0.5))

summary(m4)



m5=gls(Abund.synch.mw~wNAOwindow,data=abund.synch_mw[c(1,3,5,7,9,11,13,15,17,19,21,23,25,27,29),], cor=corAR1(0.5))
summary(m5)
```


```{r}
install.packages('orcutt')
library(orcutt)

orc.m3=
cochrane.orcutt(m3)
```

# create simple vector of labels for time window skipping every second value
```{r}
time.window5
time.window5[seq(2, length(time.window5), 2)]
time.window5_skip=time.window5

time.window5_skip[seq(2, length(time.window5), 2)]=""
```


# Plot the abundance mw over time along with NAO window
```{r}


abund.synch_mw %>% 
  ggplot()+aes(row.names(abund.synch_mw), Abund.synch.mw, group=1)+geom_point(size=3)+
  geom_path(aes(row.names(abund.synch_mw) ,wNAOwindow), col='blue')+theme(axis.text.x = element_text(angle=45, hjust=1))+
  xlab(NULL)+ylab('Mean abundance snchrony')+scale_y_continuous(sec.axis = sec_axis(~.*1, name = "Winter NAO mw"))+theme(axis.title.y.right = element_text(color = "blue")) +theme(axis.text.y.right = element_text(color='blue'))+theme(axis.title = element_text(size=12), axis.text.x = element_text(size=12)) +scale_x_discrete(labels=time.window5_skip)

```



# Run the PCoA on imputed spp data
```{r}
pcoa.imp=
log(LB10.imp[,-c(1,2)]+1) %>% 
  vegdist('bray') %>% 
  pcoa()
```

```{r}
pcoa.imp.ax=
  cbind(LB10.imp[,c(1,2)], pcoa.imp$vectors[,c(1:4)])
```

# Plot PCoA imputed spp for LB10
```{r}
pcoa.imp.ax%>% 
  mutate(year=as.numeric(year)) %>%
  #dplyr::filter(!Code %in% c('CI3', 'CI6','GI1','GI2')) %>% 
  ggplot()+aes(Axis.1, Axis.2, col=year) +geom_point(size=0.5)+geom_path()+theme_bw()+
  facet_wrap(~Code)+scale_color_gradient(high='darkblue', low='gold3')
```

# Derive matrices with scores for ax1 and ax2, to get the synchrony spatial
```{r}
pcoa.ax1.mat=
pcoa.imp.ax[,c(1,2,3)] %>% 
  pivot_wider(id_cols = c(1,2), values_from = Axis.1, names_from = year)

pcoa.ax2.mat=
pcoa.imp.ax[,c(1,2,4)] %>% 
  pivot_wider(id_cols = c(1,2), values_from = Axis.2, names_from = year)
```


# Run the moving window of PCoA axis 1 and 2 synchrony
```{r}
pcoa.ax.synch_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=1))


rownames(pcoa.ax.synch_mw)=time.window5
colnames(pcoa.ax.synch_mw)='PCoAAx1.synch.mw'



# the richness synch as simple cor
for(y in 1: length(time.window5)){
 pcoa.ax.synch_mw[y,]= mean(
   cor(
     t(pcoa.ax1.mat[,-1][,c(y:(y+5))]), method='spearman'
     )
   )
}
```

```{r}
pcoa.ax.synch_mw$PCoAAx2.synch.mw=NA

for(y in 1: length(time.window5)){
 pcoa.ax.synch_mw[y,2]= mean(
   cor(
     t(pcoa.ax2.mat[,-1][,c(y:(y+5))]), method='spearman'
     )
   )
}

```


# Add the NAOmw to the synchorny PCoA axis mw
```{r}
pcoa.ax.synch_mw$wNAO.mw= wNAO_window$NAOwindow
```

# Evident synchrony in PCoA Axis 2 during positive NAO phases
```{r}
pcoa.ax.synch_mw %>% 
  ggplot()+aes(wNAO.mw, PCoAAx1.synch.mw)+geom_point()+geom_smooth(method='lm')

pcoa.ax.synch_mw %>% 
  ggplot()+aes(wNAO.mw, PCoAAx2.synch.mw)+geom_point()+geom_smooth(method='lm')

```




# Can derive the richness and abundance time-window for temperature preferences groups
Need to first extract timeseries specific to rchness of cold, and warm spp

#*for example, use the cold.div to derive timeseries and run the moving window approach!*
```{r}
abund.ts.cold=
  cold.div %>% 
  select(c(Code, year, abund)) %>% 
  pivot_wider(names_from = year, values_from = abund) %>% as.data.frame()

abund.ts.cold$Code=NULL

# tried this to parallel what needed to run same analysis for warm taxa
#abund.ts.cold=abund.ts.cold[-c(5,6),]

```

# derive the cold abundance mw
```{r}
abund.cold.synch_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=1))


rownames(abund.cold.synch_mw)=time.window5
colnames(abund.cold.synch_mw)='Abund.cold.synch.mw'



# the richness synch as simple cor
for(y in 1: length(time.window5)){
 abund.cold.synch_mw[y,]= mean(
   cor(
     t(abund.ts.cold[,c(y:(y+5))]), method='spearman'
     )
   )
}
```

```{r}
abund.cold.synch_mw$wNAOwindow=wNAO_window$NAOwindow[match(rownames(abund.cold.synch_mw), wNAO_window$time.window5)]
```

# Plot abundance suyncrony for cold spp vs NAO
```{r}
abund.cold.synch_mw %>% 
  ggplot()+aes(wNAOwindow, Abund.cold.synch.mw, label=row.names(abund.synch_mw))+geom_point(size=2)+geom_text_repel(size=2.5)+
  geom_smooth( col='grey50', alpha=.3, method = 'lm')+ylab('Abundance Cold - synchrony mw')+ xlab('Winter NAO mw')

summary(lm(Abund.cold.synch.mw~wNAOwindow, data=abund.cold.synch_mw))
```



# Warm spp abundance ts, and moving window synch
*must delete one stream that has no variance in serveral mw, and produce NA*
```{r}
abund.ts.warm=
  warm.div %>% 
  select(c(Code, year, abund)) %>% 
  pivot_wider(names_from = year, values_from = abund) %>% as.data.frame()

abund.ts.warm$Code=NULL

abund.ts.warm=abund.ts.warm[-c(6),]

```

# derive the warm abundance mw
```{r}
abund.warm.synch_mw=as.data.frame(matrix(nrow=length(time.window5), ncol=1))


rownames(abund.warm.synch_mw)=time.window5
colnames(abund.warm.synch_mw)='Abund.warm.synch.mw'



# the richness synch as simple cor
for(y in 1: length(time.window5)){
 abund.warm.synch_mw[y,]= mean(
   cor(
     t(abund.ts.warm[,c(y:(y+5))]), method='spearman'
     )
   )
}
```

```{r}
abund.warm.synch_mw$wNAOwindow=wNAO_window$NAOwindow[match(rownames(abund.warm.synch_mw), wNAO_window$time.window5)]
```

# plot of sycnhrony mw for warm taxa
```{r}
abund.warm.synch_mw %>% 
  ggplot()+aes(wNAOwindow, Abund.warm.synch.mw, label=row.names(abund.synch_mw))+geom_point(size=2)+geom_text_repel(size=2.5)+
  geom_smooth( col='grey50', alpha=.3, method='lm')+ylab('Abundance warm - synchrony mw')+ xlab('Winter NAO mw')

summary(lm(Abund.warm.synch.mw~wNAOwindow, data=abund.warm.synch_mw))
```

*# can derive time series for TBI metrics to get synch moving window there too*
```{r}
abund.ts
TBI.loss.pa
```
# Create time window vector for the TBI metrics
```{r}
time.window.tbi=matrix(nrow=29)
for(i in 1:29){
time.window.tbi[i,]=paste(deltamat[i], deltamat[i+4], sep='_')  
}
```


# derive synch mw for TBI loss
```{r}
tbi.loss.synch_mw=as.data.frame(matrix(nrow=length(time.window.tbi), ncol=1))


rownames(tbi.loss.synch_mw)=time.window.tbi
colnames(tbi.loss.synch_mw)='Extinction.synch.mw'



# the richness synch as simple cor
for(y in 1: length(time.window.tbi)){
 tbi.loss.synch_mw[y,]= mean(
   cor(
     t(TBI.loss.pa[,c(y:(y+4))]), method='spearman'
     )
   )
}
```

```{r}
tbi.loss.synch_mw$wNAOwindow=wNAO_window$NAOwindow
#tbi.loss.synch_mw$wNAOwindo=NULL
```

# Plot of extinction synch mw vs wNAO
```{r}
tbi.loss.synch_mw %>% 
  ggplot()+aes(wNAOwindow, Extinction.synch.mw, label=row.names(tbi.loss.synch_mw))+geom_point(size=2)+geom_text_repel(size=2.5)+
  geom_smooth( col='grey50', alpha=.3, method='lm')+ylab('Extinction - synchrony mw')+ xlab('Winter NAO mw')
```

# derive synch mw for TBI gain
```{r}
# 

tbi.imm.synch_mw=as.data.frame(matrix(nrow=length(time.window.tbi), ncol=1))


rownames(tbi.imm.synch_mw)=time.window.tbi
colnames(tbi.imm.synch_mw)='Colonisation.synch.mw'



# the richness synch as simple cor
for(y in 1: length(time.window.tbi)){
 tbi.imm.synch_mw[y,]= mean(
   cor(
     t(TBI.gain.pa[,c(y:(y+4))]), method='spearman'
     )
   )
}
```




```{r}
tbi.imm.synch_mw$wNAOwindow=wNAO_window$NAOwindow
```


```{r}
tbi.imm.synch_mw %>% 
  ggplot()+aes(wNAOwindow, Colonisation.synch.mw, label=row.names(tbi.imm.synch_mw))+geom_point(size=2)+geom_text_repel(size=2.5)+
  geom_smooth( col='grey50', alpha=.3, method='lm')+ylab('Colonisation - synchrony mw')+ xlab('Winter NAO mw')
```


### Can derive the *moving window of spatial beta div*, across all streams
# Here I use the *array.beta* object created in Bio.Rmd, which holds the betadiv matrix for each year

```{r}

betalist.mw=list()# list to hold betadiv mw

# mean betadiv (across sites) for the moving window
for(y in 1: length(time.window5)){
  betalist.mw[[y]]=iCAMP::dist.3col(as.dist(apply(array.beta[,,c(y:(y+5))], c(1,2), mean)))
}

# add the time-window that match
for(y in 1: length(time.window5)){
  betalist.mw[[y]]$twin=time.window5[y]
}

# Extract and rbind all the windows for betadiversity
spatial.betadiv.mw=
do.call(rbind, betalist.mw)

# add the site pair and time window info
spatial.betadiv.mw$site.pair=paste(spatial.betadiv.mw$name1, spatial.betadiv.mw$name2, sep=":")
spatial.betadiv.mw$site.pair_tw=paste(spatial.betadiv.mw$site.pair, spatial.betadiv.mw$twin, sep="_")

```

# add site pair and timewindo info into the mean population synch moving window
```{r}
meanpop_synchronies_mw$site.pair_tw=paste(meanpop_synchronies_mw$site.pair, meanpop_synchronies_mw$timewindow, sep="_")
```

```{r}
meanpop_synchronies_mw$mean.spat.betadiv=spatial.betadiv.mw$dis[match(meanpop_synchronies_mw$site.pair_tw,
                                                                      spatial.betadiv.mw$site.pair_tw)]
```

```{r fig.width=5.5}
meanpop_synchronies_mw %>% 
  ggplot()+aes(mean.spat.betadiv, mean.pop.synch, col=NAOmw.q)+geom_smooth(method='lm', se=F)+scale_color_brewer (palette='Reds')+
  ylab('Mean population synchrony (mw)')+xlab('Mean spatial betadiversity')
```
# no apprent effect of mean spatial beta on the wangs components

```{r}

cbind.data.frame(wang_mw,
spatial.betadiv.mw %>% 
  group_by(twin) %>% 
  summarise(mean.spatbeta=mean(dis))) %>% 
  pivot_longer(cols=c(1:8)) %>% 
  ggplot()+aes(mean.spatbeta, value)+geom_point()+geom_smooth(method='lm', se=T)+facet_wrap(~name)
```



```{r}
comm.stab.mw
synch.loreau.mw
wang_mw
```







```{r}
install.packages('tsvr')
library(tsvr)

myarray[,,1]
dim(myarray)
```

```{r}

tmp1=myarray[,,1]
rowSums(tmp1)

tmp1=tmp1[rowSums(tmp1)>0,]

tsvr1=
tsvreq_classic(tmp1)

plot(tsvr1$tsvr~tsvr1$ts, type='l')
plot(tsvr1)
```

